### 2012-09-27 doc/scheme example.tex
Adding tex file for working example

\documentclass[11pt, letterpaper, oneside]{article}
\usepackage{enumitem}
\usepackage{calc}
\usepackage{ amssymb }

\begin{document}
\section{Step 1: Generate Keys}

	genkey procedure generates a secret and public key.  The public key includes vectors \textbf{L} and \textbf{R} as well as $q$ and $\mathcal{U}$.

	Questions on this step:
	\begin{enumerate}
	\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
	\end{enumerate}

\section{Step 2: Initialization}

	initialize takes the secret key and the public key generated in step 1, and an empty dictionary over the universe of elements and produces

	\begin{enumerate}
	\item a table \textbf{T} with N entries (0 to N-1) to represent the Merkle tree leaves
	\item initial state $d_{0} = \textbf{0} \in Z_{q}^{k}$ representing the root digest
	\item a label for each internal node: $\lambda(v)$
	\end{enumerate}

	Questions on this step:
	\begin{enumerate}
	\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
	\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
	\end{enumerate}

\end{document}

### 2012-10-02 doc/scheme example.tex
Summary and questions on updateVerifier and updateProver

@@
 \begin{document}
-\section{Step 1: Generate Keys}
+\section{Algortihm 1: Generate Keys}
 
@@
 
-	Questions on this step:
-	\begin{enumerate}
-	\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
-	\end{enumerate}
+%Questions on this step:
+%	\begin{enumerate}
+%	\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
+%	\end{enumerate}
 
-\section{Step 2: Initialization}
+\section{Algorithm 2: Initialization}
+\label{initialize}
 
@@
 
-	Questions on this step:
+%	Questions on this step:
+%	\begin{enumerate}
+%	\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
+%	\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
+%	\end{enumerate}
+
+\section{Algorithm 3: updateVerifier}
+\subsection{Inputs}
 	\begin{enumerate}
-	\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
-	\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
+	\item $x$ where $x \in \mathcal{U}$ is the current element of the stream
+	\item $d_{h}$ current state of $auth(D_{h})$
+	\item sk and pk - secret and public key \\
+	\end{enumerate}
+
+\subsection{Outputs}
+	\begin{enumerate}
+	\item $d_{h + 1}$ updated state of $auth(D_{h + 1})$
+	\end{enumerate}
+
+	\begin{equation}
+		d_{h+1} = d_{h} + \mathcal{D}_{r}(x, \textbf{1})
+	\end{equation}
+
+\subsection{Example}
+
+	\begin{itemize}
+	\item $\mathcal{U} = \{ 0, 1, 2, 3, 4, 5, 6, 7\}$ similar to example in figure 7.1
+	\item Suppose $x = 3$
+	\end{itemize}
+
+	Calculating $d_{h + 1}$ reduces to calculating 	$ \mathcal{D}_{r}(x, \textbf{1})$ because $d_{h}$ is known.  \\
+
+	In our case, we'd like to calculate $\mathcal{D}_{r}(3, \textbf{1})$, i.e. the digest of the root node with respect 3. \\
+
+	By definition 10, 
+	\begin{equation}
+	\mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b}(\mathcal{D}_{v_{11}}(3, \textbf{x}_{3})) 
+	\end{equation}
+where $v_{11}$ is the left child of $r$, and $\mathcal{D}_{v_{11}}$ is the partial digest of $v_{11}$.
+
+	This procedure continues for each internal node eventually reaching the leaf node. The partial digest of a leaf node, in our case, that representing the element $x = 3$ is given by:
+
+	\begin{equation}
+	\mathcal{D}_{3}(3, \textbf{1}) = \textbf{1}
+	\end{equation}
+
+	Substituting the calculated partial digests for each node in the path from the root to the updated leaf node results in:
+
+	\begin{equation}
+	\mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
+	\end{equation}
+
+	Thus, the new digest for the entire tree is given by:
+	\begin{equation}
+	d_{h+1} = d_{h} + \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
+	\end{equation}
+
+\subsection{Questions}
+\begin{enumerate}
+\item What are the roles of the secret and public key as inputs?
+In the previous email, you said the prover generates the secret key and does not output it.
+Therefore, the verifier can't have any value here for $sk$ besides null.
+\end{enumerate}
+
+\section{Algorithm 4: updateProver}
+\subsection{Input}
+	\begin{itemize}
+	\item $x$: $x \in \mathcal{U}$ is the current element of the stream
+	\item $pk$:  public key
+	\item $D_{h}$: The table, \textbf{T} which represents leaves of the Merkle tree described by the scheme
+	\item $auth(D_{h})$: The auxiliary information allowing authentication of the elements in $D_{h}$ including all labels $\lambda(v_{i})$ for internals nodes and $d_{h}$, the root digest of the whole tree.
+	\end{itemize}
+\subsection{Output}
+
+	\begin{itemize}
+	\item $D_{h + 1}$: Updated table \textbf{T}
+	\item $auth(D_{h + 1})$: Updated labels and root digest
+	\end{itemize}
+
+\subsection{Example}
+
+Using the same scenario as the verifier example where $x = 3$, the updated labels is given by:
+
+\begin{equation}
+\lambda(v_{i}) = \lambda(v_{i}) + \textbf{b}( \mathcal{D}_{v_{i}}(3, \textbf{1}) ) \textrm{ for } i = \ell, \ell -1, \ldots, 1  
+\end{equation}
+
+Calculating each $\lambda(v_{i})$ is straight forward.  In the last example, we understood how to calculate a partial digest, and the $\textbf{b}()$ function is well-defined.
+
+\subsection{Questions}
+
+From the algorithm $updateProver$ on page 10,
+
+\begin{quote}
+The algorithm sets $\textbf{T}[x]$ = $\textbf{T}[x] + \textbf{1}$, outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$. Let now
+$v_{\ell}$, $\ldots$, $v_{1}$ be the path in the lattice-based Merkle tree $T$ from node $v_{\ell}$ ($v_{\ell}$ stores the value  $\textbf{T}[x]$) to the child $v_{1}$ of the root $r$ of $T$.
+\end{quote}
+
+\begin{enumerate}
+\item How is  $\ell$ calculated in our scenario?
+
+\item What is meant by ``$v_{\ell}$ stores the value $\textbf{T}[x]$''?
+Is $v_{\ell}$ the leaf node of the Merkle tree?
+If so, equation 5.5 would update a label for $v_{\ell}$, but that contradicts our understanding of the $initialize$ algorithm which assigns labels for only internal nodes.
+
+\item Looking at the tree in Figure 7.1, we think the path is ($v_{22}$, $v_{11}$) where $\ell = 2$, because these are the only internal nodes for which there is a label to update.
+(The root has a digest, $d_{h}$, and the leaf node has an entry in $\textbf{T}$.)  Additionally, the algorithm states the path is from $v_{\ell}$ to ``the child $v_{1}$ of the root $r$ of $T$" suggesting that the last element of the path in question is the child of the root node, not the root node itself.
+Is this correct?
+
+\item A notation question about the statement ``outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$".
+Isn't \textbf{T} the whole table, whereas $\textbf{T}[x]$ is an element in \textbf{T}.
+If so, is there not a type mismatch in that statement, where $D_{h + 1}$ is the whole data structure and $\textbf{T}[x]$ is just one element?
+
+\item When/where does the root digest get updated (especially if our understanding of the path in the questions 2 and 3 is correct, i.e.
+the root is not included in that path)?
+Shouldn't there be a similar update for the root or the root digest?
+
 \end{enumerate}

### 2012-10-23 lib/sads.rb
Adding a comment to test the git package for sublime

@@ -5,4 +5,6 @@ class Sads
 	attr_accessor :labels
 	attr_reader :L, :R, :pubkey
 
+	#ADding a comment
+
 end

### 2012-10-24 doc/milestone_draft1.tex
Milestone Draft 1

--- 2012-10-02 doc/scheme example.tex
@@
 
-\subsection{Questions}
-\begin{enumerate}
-\item What are the roles of the secret and public key as inputs?
+%\subsection{Questions}
+%\begin{enumerate}
+%\item What are the roles of the secret and public key as inputs?
 In the previous email, you said the prover generates the secret key and does not output it.
 Therefore, the verifier can't have any value here for $sk$ besides null.
-\end{enumerate}
+%\end{enumerate}
 
@@
 
-\subsection{Questions}
+%\subsection{Questions}
 
-From the algorithm $updateProver$ on page 10,
+%From the algorithm $updateProver$ on page 10,
 
-\begin{quote}
-The algorithm sets $\textbf{T}[x]$ = $\textbf{T}[x] + \textbf{1}$, outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$. Let now
-$v_{\ell}$, $\ldots$, $v_{1}$ be the path in the lattice-based Merkle tree $T$ from node $v_{\ell}$ ($v_{\ell}$ stores the value  $\textbf{T}[x]$) to the child $v_{1}$ of the root $r$ of $T$.
-\end{quote}
+%\begin{quote}
+%The algorithm sets $\textbf{T}[x]$ = $\textbf{T}[x] + \textbf{1}$, outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$. Let now
+%$v_{\ell}$, $\ldots$, $v_{1}$ be the path in the lattice-based Merkle tree $T$ from node $v_{\ell}$ ($v_{\ell}$ stores the value  $\textbf{T}[x]$) to the child $v_{1}$ of the root $r$ of $T$.
+%\end{quote}
 
-\begin{enumerate}
-\item How is  $\ell$ calculated in our scenario?
+%\begin{enumerate}
+%\item How is  $\ell$ calculated in our scenario?
 
-\item What is meant by ``$v_{\ell}$ stores the value $\textbf{T}[x]$''?
+%\item What is meant by ``$v_{\ell}$ stores the value $\textbf{T}[x]$''?
 Is $v_{\ell}$ the leaf node of the Merkle tree?
@@
 
-\item Looking at the tree in Figure 7.1, we think the path is ($v_{22}$, $v_{11}$) where $\ell = 2$, because these are the only internal nodes for which there is a label to update.
+%\item Looking at the tree in Figure 7.1, we think the path is ($v_{22}$, $v_{11}$) where $\ell = 2$, because these are the only internal nodes for which there is a label to update.
 (The root has a digest, $d_{h}$, and the leaf node has an entry in $\textbf{T}$.)  Additionally, the algorithm states the path is from $v_{\ell}$ to ``the child $v_{1}$ of the root $r$ of $T$" suggesting that the last element of the path in question is the child of the root node, not the root node itself.
@@
 
-\item A notation question about the statement ``outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$".
+%\item A notation question about the statement ``outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$".
 Isn't \textbf{T} the whole table, whereas $\textbf{T}[x]$ is an element in \textbf{T}.
@@
 
-\item When/where does the root digest get updated (especially if our understanding of the path in the questions 2 and 3 is correct, i.e.
+%\item When/where does the root digest get updated (especially if our understanding of the path in the questions 2 and 3 is correct, i.e.
 the root is not included in that path)?
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Implementation Issues\\}
+\subsection{Proof of Concept}
+	\begin{enumerate}
+	\item In order to implement a streaming authenticated data structure based on lattices, a fast
+	arithmetic library for matrices and vector is required.
+	\item Matrix multiplication mod \texttt{q} ( \texttt{q poly(k)}) is needed.
 	\end{enumerate}
 
+\subsection{Applicable Libraries\\}
+	\begin{enumerate}
+	\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length
+	integers, for vectors, matrices and polynomials over the integers and over the finite fields and for 
+	arbitrary precision floating point arithmetic. \texttt{http://shoup.net/ntl/}
+	\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of
+the important concrete classes of 
+	structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, 
+	module theory and the theory of algebras. In addition, certain
+families of structures from algebraic 
+	geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
+	\end{enumerate}
+
+\subsection{ Library Prototype\\}
+
+	We decided to start a feasible implementation of the model by trying to find the right library
+	for matrix operations and arbitrary precision floating point arithmetics. In our investigations,
+	we had the following observations :
+
+	\begin{itemize}
+
+	\item It seems that there has been more recent updates and user discussion links regarding MAGMA
+	and the library has attracted more draw in terms of being more up to dated and user-friendly.
+	\item MAGMA is a collection of linear algebra GPU accelerated libraries designed and implemented
+	by the team that developed LAPACK. Since we were not completely sure about MAGMA's dependency on
+	CUDA and other related libraries we did not go into further detail at this stage.
+	\item We decided to work with Ruby (or Python) for the first implementation since they both have a
+	standard library documentation for matrices.
+	\item Starting with this idea, we can smoothly find the best matching library for our system 
+	model which helps us to know exactly what requirements we need and which library matches these 
+	requirements accordingly.
+	\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\end{enumerate}
+
+\section{Matrix Multiplication in Ruby}
+
+\subsection{The \texttt{[row $\times$ column]} multiplication:}
+	\begin{itemize}
+	\item $ C = A_{n \times n} \times B_{n \times n}$
+	\item where $ c_{ij} = \sum\limits_{k=1}^n a_{ik} . b_{kj}   $
+	\item complexity : $\Theta(n^3)$ \\
+	\end{itemize}
+
+\subsection{Divide and Conquer}
+
+	\begin{itemize}
+
+        \item Each of the matrices is divided into four $\frac{n}{2} \times \frac{n}{2}$
+	\item The multiplication of two $ n \times  n $ matrices is reduced to 8 multiplications of 
+	two $\frac{n}{2} \times \frac{n}{2}$ matrices and four $\frac{n}{2} \times \frac{n}{2}$ matrix additions.
+	\item recurrence : $T(n) = 8.T. \frac{n}{2} + \Theta(n^2) \rightarrow \Theta(n^3)$ \\
+	\end{itemize}
+
+\subsection{Strassen's Algorithm}
+
+	\begin{itemize}
+
+        \item Bulky for implementation and not numerically stable
+        \item Useful and advantageous for large matrices
+	\item $ \Theta(n^{log7}) \approx \Theta(n^{2.807})$ \\
+	\end{itemize}
+
+\subsection{Coppersmith Winograd Algorithm}
+
+	\begin{itemize}
+
+        \item Theoretical optimum; never used in practice.
+	\item $ \Theta(n^{2.376}) $ \\
+	\end{itemize}
+
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2012-10-24 lib/sads.rb
Removing example comment

@@ -5,6 +5,4 @@ class Sads
 	attr_accessor :labels
 	attr_reader :L, :R, :pubkey
 
-	#ADding a comment
-
 end

### 2012-10-24 lib/sads.rb
Adding method stubs

@@ -2,7 +2,27 @@
 
 class Sads
 	
-	attr_accessor :labels
+	attr_accessor :labels, :leaves, :digest
 	attr_reader :L, :R, :pubkey
+
+	def initialize()
+		# Size of Universe
+		# 
+
+
+	end
+
+
+	def addElement(ele)
+
+	end
+
+	def removeElement(ele)
+
+	end
+
+	def exists?(ele)
+		return false
+	end
 
 end

### 2012-10-24 spec/sads_spec.rb
Some pending tests to define interface

@@ -13,7 +13,6 @@ describe Sads do
 	end
 
 	describe "instance fields" do
-		# Put feature tests here
 
 		it "should have a structure called labels" do
 			@sad.should respond_to(:labels)
@@ -27,8 +26,73 @@ describe Sads do
 		it "should have a public and secret keypair" do
 			@sad.should respond_to(:pubkey)
 		end
+
+		it "should have a field called leaves" do
+			@sad.should respond_to(:leaves)
+		end
+
+		it "should have a field called digest" do
+			@sad.should respond_to(:digest)
+		end
 	end # describe instance fields
 
+	describe "initialize" do
 
-	
+		it "should initialize labels as a map" do
+			@sad.labels.should be_empty
+		end
+
+		it "should initialize leaves as a map" do
+			pending
+		end
+
+		it "should initialize digest as a SOMETHING" do
+			pending
+		end
+
+	end # describe initialize
+
+	describe "add element" do
+
+		it "should update leaves table if the element is new" do
+			pending
+		end
+
+		it "should update the labels of the internal nodes on the path to the root" do
+			pending
+		end
+
+		it "should update the digest of the tree" do
+			pending
+		end
+
+	end # describe add element
+
+	describe "remove element" do
+
+		it "should update the leaves table if the element was in the tree" do
+			pending
+		end
+
+		it "should update the labels of the internal nodes on the path to the root" do
+			pending
+		end
+
+		it "should update the digest of the tree" do
+			pending
+		end
+
+	end # describe remove element
+
+	describe "exists?" do
+
+		it "should return true for elements in the tree" do
+			pending
+		end
+
+		it "should return false for elements not in the tree" do
+			pending
+		end
+
+	end # describe 
 end # describe Sads

### 2012-10-24 lib/sads.rb
Passing label init

@@ -8,8 +8,9 @@ class Sads
 	def initialize()
 		# Size of Universe
 		# 
-
-
+		@labels = {}
+		
+		
 	end
 
 

### 2012-10-24 lib/sads.rb
Passing leaves init

@@ -9,8 +9,9 @@ class Sads
 		# Size of Universe
 		# 
 		@labels = {}
-		
-		
+		@leaves = {}
+
+
 	end
 
 

### 2012-10-24 spec/sads_spec.rb

@@ -43,7 +43,7 @@ describe Sads do
 		end
 
 		it "should initialize leaves as a map" do
-			pending
+			@sad.leaves.should be_empty
 		end
 
 		it "should initialize digest as a SOMETHING" do

### 2012-10-24 doc/milestone_draft1.tex
Putting one sentence per line for easier version control

@@
 \section{Algortihm 1: Generate Keys}
-
-	genkey procedure generates a secret and public key.  The public key includes vectors \textbf{L} and \textbf{R} as well as $q$ and $\mathcal{U}$.
+genkey procedure generates a secret and public key.
+The public key includes vectors \textbf{L} and \textbf{R} as well as $q$ and $\mathcal{U}$.
 
@@
 \label{initialize}
-
 initialize takes the secret key and the public key generated in step 1, and an empty dictionary over the universe of elements and produces
@@
 
-	This procedure continues for each internal node eventually reaching the leaf node. The partial digest of a leaf node, in our case, that representing the element $x = 3$ is given by:
+This procedure continues for each internal node eventually reaching the leaf node.
+The partial digest of a leaf node, in our case, that representing the element $x = 3$ is given by:
 
@@
 	integers, for vectors, matrices and polynomials over the integers and over the finite fields and for 
-	arbitrary precision floating point arithmetic. \texttt{http://shoup.net/ntl/}
+	arbitrary precision floating point arithmetic.
+	\texttt{http://shoup.net/ntl/}
 	\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of
-the important concrete classes of 
-	structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, 
-	module theory and the theory of algebras. In addition, certain
-families of structures from algebraic 
-	geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
+the important concrete classes of structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, module theory and the theory of algebras.
+In addition, certain
+families of structures from algebraic geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
 	\end{enumerate}
@@
 
-	We decided to start a feasible implementation of the model by trying to find the right library
-	for matrix operations and arbitrary precision floating point arithmetics. In our investigations,
-	we had the following observations :
+	We decided to start a feasible implementation of the model by trying to find the right library for matrix operations and arbitrary precision floating point arithmetics.
+In our investigations, we had the following observations :
 
@@
 
-	\item It seems that there has been more recent updates and user discussion links regarding MAGMA
-	and the library has attracted more draw in terms of being more up to dated and user-friendly.
-	\item MAGMA is a collection of linear algebra GPU accelerated libraries designed and implemented
-	by the team that developed LAPACK. Since we were not completely sure about MAGMA's dependency on
-	CUDA and other related libraries we did not go into further detail at this stage.
-	\item We decided to work with Ruby (or Python) for the first implementation since they both have a
-	standard library documentation for matrices.
-	\item Starting with this idea, we can smoothly find the best matching library for our system 
-	model which helps us to know exactly what requirements we need and which library matches these 
-	requirements accordingly.
+	\item It seems that there has been more recent updates and user discussion links regarding MAGMA and the library has attracted more draw in terms of being more up to dated and user-friendly.
+	\item MAGMA is a collection of linear algebra GPU accelerated libraries designed and implemented by the team that developed LAPACK.
+Since we were not completely sure about MAGMA's dependency on CUDA and other related libraries we did not go into further detail at this stage.
+	\item We decided to work with Ruby (or Python) for the first implementation since they both have a standard library documentation for matrices.
+	\item Starting with this idea, we can smoothly find the best matching library for our system model which helps us to know exactly what requirements we need and which library matches these requirements accordingly.
 	\end{itemize}

### 2012-10-25 spec/sads_spec.rb
Some specs

@@ -47,6 +47,7 @@ describe Sads do
 		end
 
 		it "should initialize digest as a SOMETHING" do
+			# @sad.digest.should_not be_nil
 			pending
 		end
 
@@ -55,7 +56,7 @@ describe Sads do
 	describe "add element" do
 
 		it "should update leaves table if the element is new" do
-			pending
+			@sad.leaves.should_not be_empty
 		end
 
 		it "should update the labels of the internal nodes on the path to the root" do

### 2012-10-25 lib/sads.rb
Somenotes

@@ -6,12 +6,12 @@ class Sads
 	attr_reader :L, :R, :pubkey
 
 	def initialize()
-		# Size of Universe
-		# 
+		# Input parameteres?
+		#  # Size of Universe
+
 		@labels = {}
 		@leaves = {}
-
-
+		# @digest = 0
 	end
 
 

### 2012-10-25 lib/sads.rb
Basic exists tests passing and some formatting

@@ -16,7 +16,7 @@ class Sads
 
 
 	def addElement(ele)
-
+		leaves[ele] = 1
 	end
 
 	def removeElement(ele)
@@ -24,7 +24,7 @@ class Sads
 	end
 
 	def exists?(ele)
-		return false
+		return leaves.include?(ele)
 	end
 
 end

### 2012-10-25 spec/sads_spec.rb

@@ -50,12 +50,12 @@ describe Sads do
 			# @sad.digest.should_not be_nil
 			pending
 		end
-
 	end # describe initialize
 
 	describe "add element" do
 
 		it "should update leaves table if the element is new" do
+			@sad.addElement(1)
 			@sad.leaves.should_not be_empty
 		end
 
@@ -66,7 +66,6 @@ describe Sads do
 		it "should update the digest of the tree" do
 			pending
 		end
-
 	end # describe add element
 
 	describe "remove element" do
@@ -82,18 +81,22 @@ describe Sads do
 		it "should update the digest of the tree" do
 			pending
 		end
-
 	end # describe remove element
 
 	describe "exists?" do
 
 		it "should return true for elements in the tree" do
-			pending
+			@sad.addElement(1)
+			@sad.exists?(1).should be_true
 		end
 
 		it "should return false for elements not in the tree" do
-			pending
+			@sad.addElement(1)
+			@sad.exists?(2).should be_false
 		end
 
+		it "should return false for empty tree" do
+			@sad.exists?(1).should be_false
+		end
 	end # describe 
 end # describe Sads

### 2012-10-25 lib/sads.rb
Add and remove update leaves hash

@@ -16,11 +16,11 @@ class Sads
 
 
 	def addElement(ele)
-		leaves[ele] = 1
+		leaves.store(ele, 1)
 	end
 
 	def removeElement(ele)
-
+		leaves.delete ele
 	end
 
 	def exists?(ele)

### 2012-10-25 spec/sads_spec.rb

@@ -71,7 +71,9 @@ describe Sads do
 	describe "remove element" do
 
 		it "should update the leaves table if the element was in the tree" do
-			pending
+			@sad.addElement(1)
+			@sad.removeElement(1)
+			@sad.exists?(1).should be_false
 		end
 
 		it "should update the labels of the internal nodes on the path to the root" do

### 2012-10-26 lib/sads.rb
Added some comments to field names and pending test explanations

@@ -3,7 +3,15 @@
 class Sads
 	
 	attr_accessor :labels, :leaves, :digest
-	attr_reader :L, :R, :pubkey
+
+	# Security Parameter
+	attr_reader :k
+
+	# Left and Right vectors for algebraic hash function
+	attr_reader :L, :R
+
+	# Encryption key pair
+	attr_reader :public_key, :secret_key
 
 	def initialize()
 		# Input parameteres?

### 2012-10-26 spec/sads_spec.rb

@@ -24,7 +24,7 @@ describe Sads do
 		end
 
 		it "should have a public and secret keypair" do
-			@sad.should respond_to(:pubkey)
+			pending "Accessible via accessor so test isn't fragile w.r.t field name"
 		end
 
 		it "should have a field called leaves" do
@@ -37,6 +37,18 @@ describe Sads do
 	end # describe instance fields
 
 	describe "initialize" do
+
+		it "should initialize L vector" do
+			pending "Needs specification"
+		end
+
+		it "should initialize R vector" do
+			pending "Needs specification"
+		end
+
+		it "should initialize the public/secret keypair" do
+			pending "Needs specification"
+		end
 
 		it "should initialize labels as a map" do
 			@sad.labels.should be_empty

### 2012-10-29 doc/milestone_draft1.tex
Milestone Draft 2

--- 2012-10-24 doc/milestone_draft1.tex
@@
 \begin{document}
-\section{Algortihm 1: Generate Keys}
+
+\title{\textbf{ CMCS 818J Project Milestone Report } \\  \small{Christopher Imbriano \& Behzad Koosha}}
+\maketitle
+
+\section{Abstract}
+
+	As it has been described in the project proposal, the aim for this project is to present and implement a new model for verifying data structure queries in a streaming scenario.
+The proposed scheme gives rise to new cryptographic constructions with improved properties such as dynamic cryptographic accumulator; parallel online memory checking and a space efficient three party authenticated data structure.
+
+\section{Introduction}
+
+	The introduced scheme, streaming authenticated data structure, does not require any interaction between the client and the server while the stream is observed.
+Three important properties is obtained using this scheme : a.
+Independence of prover and verifier : the prover and the verifier update their status independently and with no interaction.
+b.
+Efficiency : It can be shown that an exponential improvement for many commonly queries in the prover's running time can be achieved.
+c.
+Expressiveness : a wide range of queries such as membership/ non-membership, successor, range search and frequencies can be supported using this scheme.
+
+\section{Previous Work}
+
+	Most of the existing streaming verifiable protocols are efficient in terms of verifier complexity; it can be observed that in all of them the prover complexity is linear in the size of stream which this limits the applicability of the protocols.
+Other works include verifiable computation and authenticated data structures, which are not directly applicable to the streaming data set or their application results in increased complexities or interactive protocols.
+
+\section{Algorithm}
+
+	In the following, we will represent the steps that the streaming authenticated data structure follows to implement range search queries.
+
+\subsection{ 1: Generate Keys}
 
@@
 
-\section{Algorithm 2: Initialization}
+\subsection{ 2: Initialization}
 \label{initialize}
@@
 
-\section{Algorithm 3: updateVerifier}
-\subsection{Inputs}
+\subsection{ 3: updateVerifier}
+\subsubsection{Inputs}
 	\begin{enumerate}
@@
 
-\subsection{Outputs}
+\subsubsection{Outputs}
 	\begin{enumerate}
@@
 
-\subsection{Example}
+\subsubsection{Example}
 
@@
 
-\section{Algorithm 4: updateProver}
-\subsection{Input}
+\subsection{ 4: updateProver}
+\subsubsection{Input}
 	\begin{itemize}
@@
 	\end{itemize}
-\subsection{Output}
+\subsubsection{Output}
 
@@
 
+\section{Future Work}
+
+	For the remaining time in the semester, we plan to implement our idea in Ruby and then move on to implenting the structure to the right library such as MAGMA.
+We aim to design and implement the scheme for applications that can fully enable efficient verification of data structure queries on a stream.
+In addition, we will be meeting with Elaine (and Babis via Skype) to discuss and exchange thoughts about our approach to solve the problem as we have already practiced it before.
+
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-10-30 spec/sads_spec.rb
Clarified test

@@ -92,7 +92,7 @@ describe Sads do
 			pending
 		end
 
-		it "should update the digest of the tree" do
+		it "should update the root digest of the tree" do
 			pending
 		end
 	end # describe remove element

### 2012-10-30 lib/sads_util.rb
Adding sads_util for helper methods

@@ -1,0 +1,10 @@
+
+def calculate_q(n, k, w)
+	# q is the smallest prime satisfying
+	# 
+	# q / log q >= n * 2k * w( sqrt(k * log k))
+end
+
+def calculate_mu(k, q)
+	#2 * k * round_up(log q)
+end

### 2012-10-31 lib/sads.rb
Removing util in place of private methods in class

@@ -1,11 +1,9 @@
 # sads.rb
+require 'sads_util'
 
 class Sads
 	
 	attr_accessor :labels, :leaves, :digest
-
-	# Security Parameter
-	attr_reader :k
 
 	# Left and Right vectors for algebraic hash function
 	attr_reader :L, :R
@@ -13,9 +11,32 @@ class Sads
 	# Encryption key pair
 	attr_reader :public_key, :secret_key
 
+	# Security Parameter
+	attr_reader :k
+
+	# Size of Universe
+	attr_reader :universe_size_M
+
+	# Upper bound on size of stream
+	attr_reader :stream_bound_n
+
+	# Algorithm Paramters
+	# q
+	attr_reader :q
+
+	#mu
+	attr_reader :mu
+
+	#beta
+	attr_reader :beta
+
 	def initialize()
 		# Input parameteres?
 		#  # Size of Universe
+		
+		@k = 500
+		# @stream_bound_n = poly(k)
+		
 
 		@labels = {}
 		@leaves = {}
@@ -34,5 +55,22 @@ class Sads
 	def exists?(ele)
 		return leaves.include?(ele)
 	end
+
+	private do	
+
+		def calculate_q(n, k, w)
+			# q is the smallest prime satisfying
+			# 
+			# q / log (q + 1) >= n * 2k * w( sqrt(k * log k))
+		end
+
+		def calculate_mu(k, q)
+			#2 * k * round_up(log q)
+		end
+
+		def calculate_beta(n, mu)
+			n * Math.sqrt(mu)
+		end
+	end # private
 
 end

### 2012-10-31 lib/sads_util.rb

@@ -1,10 +1,0 @@
-
-def calculate_q(n, k, w)
-	# q is the smallest prime satisfying
-	# 
-	# q / log q >= n * 2k * w( sqrt(k * log k))
-end
-
-def calculate_mu(k, q)
-	#2 * k * round_up(log q)
-end

### 2012-10-31 spec/sads_spec.rb
Stub for L and R vectors

@@ -34,6 +34,14 @@ describe Sads do
 		it "should have a field called digest" do
 			@sad.should respond_to(:digest)
 		end
+
+		describe "L and R vectors" do
+
+			it "should be a vector sized k * m and have elements in Z_q" do
+				pending "Not implemented"
+			end
+
+		end #describe L and R vectors
 	end # describe instance fields
 
 	describe "initialize" do

### 2012-10-31 lib/sads.rb
Adding some basics to initialize

@@ -1,5 +1,5 @@
 # sads.rb
-require 'sads_util'
+require 'matrix'
 
 class Sads
 	
@@ -37,6 +37,27 @@ class Sads
 		@k = 500
 		# @stream_bound_n = poly(k)
 		
+		q    = calculate_q(@stream_bound_n, @k, 1)
+		mu   = calculate_mu(@k, q)
+		beta = calculate_beta(4, 9)
+
+
+		# m = mu / 2
+		m = 10
+
+		dim = @k * m
+		l_array = Array.new
+		dim.times do 
+			l_array.push rand(q) 
+		end
+
+		r_array = Array.new
+		dim.times do 
+			r_array.push rand(q) 
+		end
+
+		@L = Vector.elements(l_array)
+		@R = Vector.elements(r_array)
 
 		@labels = {}
 		@leaves = {}
@@ -56,21 +77,20 @@ class Sads
 		return leaves.include?(ele)
 	end
 
-	private do	
+	private
 
-		def calculate_q(n, k, w)
-			# q is the smallest prime satisfying
-			# 
-			# q / log (q + 1) >= n * 2k * w( sqrt(k * log k))
-		end
+	def calculate_q(n, k, w)
+		# q is the smallest prime satisfying
+		# 
+		# q / log (q + 1) >= n * 2k * w( sqrt(k * log k))
+	end
 
-		def calculate_mu(k, q)
-			#2 * k * round_up(log q)
-		end
+	def calculate_mu(k, q)
+		#2 * k * round_up(log q)
+	end
 
-		def calculate_beta(n, mu)
-			n * Math.sqrt(mu)
-		end
-	end # private
-
+	def calculate_beta(n, mu)
+		n * Math.sqrt(mu)
+	end
+	# end private
 end

### 2012-10-31 spec/sads_spec.rb

@@ -47,7 +47,8 @@ describe Sads do
 	describe "initialize" do
 
 		it "should initialize L vector" do
-			pending "Needs specification"
+			expect(@sad.L.size).to eq(10)
+			# pending "Needs specification"
 		end
 
 		it "should initialize R vector" do

### 2012-10-31 tc_sads.rb
Unit test file, at top level until we figure out correct layout

@@ -1,0 +1,97 @@
+require	'./sads'
+require 'test/unit'
+
+class TestSads < Test::Unit::TestCase
+	def setup
+		@sad = Sads.new
+	end
+
+	def teardown
+		#Nothing yet
+	end
+
+	def test_constructor
+		assert_not_nil(@sad)
+	end
+
+
+	# Leave test here: failing test at the moment
+	def test_calc_mu_q_0
+		assert(false, "Not implemented")
+		assert_equal(0, @sad.calculate_mu(0,0))
+	end
+
+	def test_mu_is_not_a_float
+		assert(false, "Not implemented")
+	end
+	
+	def test_calc_q
+		assert(false, "Not implemented")
+	end 
+	
+
+	def test_calc_mu
+		assert_equal(12, @sad.calculate_mu(2,8))
+		assert_equal(16, @sad.calculate_mu(2,9))
+	end
+
+
+	def test_calc_beta
+		assert_equal(6, @sad.calculate_beta(2,9))
+		assert_in_delta(8.94, @sad.calculate_beta(2,20), 0.01)
+	end
+
+	def test_init_L_R_are_matricecs
+		assert_instance_of(Matrix, @sad.L)
+		assert_instance_of(Matrix, @sad.R)
+	end
+
+	def test_init_L_R_are_correct_size
+		assert_equal( @sad.L.row_size, @sad.k)
+		assert_equal( @sad.L.column_size, @sad.mu / 2)
+		assert_equal( @sad.R.row_size, @sad.k)
+		assert_equal( @sad.R.column_size, @sad.mu / 2)
+	end
+
+	def test_init_L_R_are_mod_q
+		@sad.L.each do |item|
+			assert(item < @sad.q)
+		end
+
+		@sad.R.each do |item|
+			assert(item < @sad.q)
+		end
+	end
+
+	def test_init_L_R_are_not_equal
+		assert( false == @sad.L.eql?(@sad.R))
+	end
+
+	def test_hash_elements_are_mod_q
+		@sad.hash(x,y).each do |item|
+			assert(item < @sad.q)
+		end
+	end
+
+	def test_hash_result_is_a_matrix
+		assert_instance_of(Matrix, @sad.hash(x,y))
+	end
+end
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2012-10-31 sads.rb
Unit testing basic initializations of Sads

@@ -33,35 +33,13 @@ class Sads
 	def initialize()
 		# Input parameteres?
 		#  # Size of Universe
+				
+		@k = 5
+		@q = 17
+		@mu   = calculate_mu(@k, @q)
 		
-		@k = 500
-		# @stream_bound_n = poly(k)
-		
-		q    = calculate_q(@stream_bound_n, @k, 1)
-		mu   = calculate_mu(@k, q)
-		beta = calculate_beta(4, 9)
+		init_L_R
 
-
-		# m = mu / 2
-		m = 10
-
-		dim = @k * m
-		l_array = Array.new
-		dim.times do 
-			l_array.push rand(q) 
-		end
-
-		r_array = Array.new
-		dim.times do 
-			r_array.push rand(q) 
-		end
-
-		@L = Vector.elements(l_array)
-		@R = Vector.elements(r_array)
-
-		@labels = {}
-		@leaves = {}
-		# @digest = 0
 	end
 
 
@@ -77,20 +55,32 @@ class Sads
 		return leaves.include?(ele)
 	end
 
-	private
+	# private below here
 
-	def calculate_q(n, k, w)
+	def calculate_q(n, k)
 		# q is the smallest prime satisfying
 		# 
 		# q / log (q + 1) >= n * 2k * w( sqrt(k * log k))
+		
+		# Temporarily Set to arbirary value
+		n * k
 	end
 
 	def calculate_mu(k, q)
-		#2 * k * round_up(log q)
+		2 * k * Math.log2(q).ceil
 	end
 
 	def calculate_beta(n, mu)
 		n * Math.sqrt(mu)
+	end
+
+	def init_L_R
+		@L = Matrix.build(@k, @mu / 2) { rand @q }
+		@R = Matrix.build(@k, @mu / 2) { rand @q }
+	end
+
+	def hash(x, y)
+		@L * x + @R * y
 	end
 	# end private
 end

### 2012-10-31 spec/sads_spec.rb
Commenting out most of spec.  This should be used for testing the public interface

@@ -6,120 +6,120 @@ describe Sads do
 	it "should be a sads object" do
 		Sads.new.should be_a(Sads)
 	end
-	
+
 	before(:each) do
 		# Pre-test setup
 		@sad = Sads.new()
 	end
 
-	describe "instance fields" do
+	# describe "instance fields" do
 
-		it "should have a structure called labels" do
-			@sad.should respond_to(:labels)
-		end
+	# 	it "should have a structure called labels" do
+	# 		@sad.should respond_to(:labels)
+	# 	end
 
-		it "should have L (left) and R (right) matrices (or column vectors)" do
-			@sad.should respond_to(:L)
-			@sad.should respond_to(:R)
-		end
+	# 	it "should have L (left) and R (right) matrices (or column vectors)" do
+	# 		@sad.should respond_to(:L)
+	# 		@sad.should respond_to(:R)
+	# 	end
 
-		it "should have a public and secret keypair" do
-			pending "Accessible via accessor so test isn't fragile w.r.t field name"
-		end
+	# 	it "should have a public and secret keypair" do
+	# 		pending "Accessible via accessor so test isn't fragile w.r.t field name"
+	# 	end
 
-		it "should have a field called leaves" do
-			@sad.should respond_to(:leaves)
-		end
+	# 	it "should have a field called leaves" do
+	# 		@sad.should respond_to(:leaves)
+	# 	end
 
-		it "should have a field called digest" do
-			@sad.should respond_to(:digest)
-		end
+	# 	it "should have a field called digest" do
+	# 		@sad.should respond_to(:digest)
+	# 	end
 
-		describe "L and R vectors" do
+	# 	describe "L and R vectors" do
 
-			it "should be a vector sized k * m and have elements in Z_q" do
-				pending "Not implemented"
-			end
+	# 		it "should be a vector sized k * m and have elements in Z_q" do
+	# 			pending "Not implemented"
+	# 		end
 
-		end #describe L and R vectors
-	end # describe instance fields
+	# 	end #describe L and R vectors
+	# end # describe instance fields
 
-	describe "initialize" do
+	# describe "initialize" do
 
-		it "should initialize L vector" do
-			expect(@sad.L.size).to eq(10)
-			# pending "Needs specification"
-		end
+	# 	it "should initialize L vector" do
+	# 		expect(@sad.L.size).to eq(10)
+	# 		# pending "Needs specification"
+	# 	end
 
-		it "should initialize R vector" do
-			pending "Needs specification"
-		end
[…]

### 2012-11-01 test/helper.rb
Using jeweler to manage project

@@ -1,0 +1,18 @@
+require 'rubygems'
+require 'bundler'
+begin
+  Bundler.setup(:default, :development)
+rescue Bundler::BundlerError => e
+  $stderr.puts e.message
+  $stderr.puts "Run `bundle install` to install missing gems"
+  exit e.status_code
+end
+require 'test/unit'
+require 'shoulda'
+
+$LOAD_PATH.unshift(File.join(File.dirname(__FILE__), '..', 'lib'))
+$LOAD_PATH.unshift(File.dirname(__FILE__))
+require 'sads'
+
+class Test::Unit::TestCase
+end

### 2012-11-01 test/test_sads.rb

@@ -1,0 +1,74 @@
+require 'helper'
+
+class TestSads < Test::Unit::TestCase
+	def setup
+		@sad = Sads.new
+	end
+
+	def test_constructor
+		assert_not_nil(@sad)
+	end
+
+	# Leave test here: failing test at the moment
+	def test_calc_mu_q_0
+		assert(false, "Not implemented")
+		assert_equal(0, @sad.calculate_mu(0,0))
+	end
+
+	def test_mu_is_not_a_float
+		assert(false, "Not implemented")
+	end
+	
+	def test_calc_q
+		assert(false, "Not implemented")
+	end 
+	
+
+	def test_calc_mu
+		assert_equal(12, @sad.calculate_mu(2,8))
+		assert_equal(16, @sad.calculate_mu(2,9))
+	end
+
+
+	def test_calc_beta
+		assert_equal(6, @sad.calculate_beta(2,9))
+		assert_in_delta(8.94, @sad.calculate_beta(2,20), 0.01)
+	end
+
+	def test_init_L_R_are_matricecs
+		assert_instance_of(Matrix, @sad.L)
+		assert_instance_of(Matrix, @sad.R)
+	end
+
+	def test_init_L_R_are_correct_size
+		assert_equal( @sad.L.row_size, @sad.k)
+		assert_equal( @sad.L.column_size, @sad.mu / 2)
+		assert_equal( @sad.R.row_size, @sad.k)
+		assert_equal( @sad.R.column_size, @sad.mu / 2)
+	end
+
+	def test_init_L_R_are_mod_q
+		@sad.L.each do |item|
+			assert(item < @sad.q)
+		end
+
+		@sad.R.each do |item|
+			assert(item < @sad.q)
+		end
+	end
+
+	def test_init_L_R_are_not_equal
+		assert( false == @sad.L.eql?(@sad.R))
+	end
+
+	def test_hash_elements_are_mod_q
+		@sad.hash(x,y).each do |item|
+			assert(item < @sad.q)
+		end
+	end
+
+	def test_hash_result_is_a_matrix
+		assert_instance_of(Matrix, @sad.hash(x,y))
+	end
+
+end

### 2012-11-01 tc_sads.rb
Removing old test file

@@ -1,97 +1,0 @@
-require	'./sads'
-require 'test/unit'
-
-class TestSads < Test::Unit::TestCase
-	def setup
-		@sad = Sads.new
-	end
-
-	def teardown
-		#Nothing yet
-	end
-
-	def test_constructor
-		assert_not_nil(@sad)
-	end
-
-
-	# Leave test here: failing test at the moment
-	def test_calc_mu_q_0
-		assert(false, "Not implemented")
-		assert_equal(0, @sad.calculate_mu(0,0))
-	end
-
-	def test_mu_is_not_a_float
-		assert(false, "Not implemented")
-	end
-	
-	def test_calc_q
-		assert(false, "Not implemented")
-	end 
-	
-
-	def test_calc_mu
-		assert_equal(12, @sad.calculate_mu(2,8))
-		assert_equal(16, @sad.calculate_mu(2,9))
-	end
-
-
-	def test_calc_beta
-		assert_equal(6, @sad.calculate_beta(2,9))
-		assert_in_delta(8.94, @sad.calculate_beta(2,20), 0.01)
-	end
-
-	def test_init_L_R_are_matricecs
-		assert_instance_of(Matrix, @sad.L)
-		assert_instance_of(Matrix, @sad.R)
-	end
-
-	def test_init_L_R_are_correct_size
-		assert_equal( @sad.L.row_size, @sad.k)
-		assert_equal( @sad.L.column_size, @sad.mu / 2)
-		assert_equal( @sad.R.row_size, @sad.k)
-		assert_equal( @sad.R.column_size, @sad.mu / 2)
-	end
-
-	def test_init_L_R_are_mod_q
-		@sad.L.each do |item|
-			assert(item < @sad.q)
-		end
-
-		@sad.R.each do |item|
-			assert(item < @sad.q)
-		end
-	end
-
-	def test_init_L_R_are_not_equal
-		assert( false == @sad.L.eql?(@sad.R))
-	end
-
-	def test_hash_elements_are_mod_q
-		@sad.hash(x,y).each do |item|
-			assert(item < @sad.q)
-		end
-	end
-
-	def test_hash_result_is_a_matrix
-		assert_instance_of(Matrix, @sad.hash(x,y))
-	end
-end
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2012-11-01 test/helper.rb
Added simplecov gem fo code coverage, ignoring its generated file. Added coverage rake task

@@ -1,3 +1,6 @@
+require 'simplecov'
+SimpleCov.start if ENV['COVERAGE']
+
 require 'rubygems'
 require 'bundler'
 begin

### 2012-11-01 test/helper.rb
Only need require 'simplecov' due to centralized configuration

@@ -1,5 +1,5 @@
 require 'simplecov'
-SimpleCov.start if ENV['COVERAGE']
+#SimpleCov.start if ENV['COVERAGE']
 
 require 'rubygems'
 require 'bundler'

### 2012-11-02 doc/milestone_draft1.tex
Milestone Draft Thursday

@@
 
-\section{Algorithm}
+\section{Our Approach}
 
-	In the following, we will represent the steps that the streaming authenticated data structure follows to implement range search queries.
+%In the following, we will represent the steps that the streaming authenticated data structure follows to implement range search queries.
+Taking into account that the bottomline of our scheme is the Merkle tree structure and using lattice based hash
+function with homomorphic property, we are implementing the algorithm in Ruby language at this stage and 
+defining classes for actually testing and configuring different steps in the algorithm. \\
 
-\subsection{ 1: Generate Keys}
+Languages such as Python or Ruby support
 
-	genkey procedure generates a secret and public key.  The public key includes vectors \textbf{L} and \textbf{R} as well as $q$ and $\mathcal{U}$.
+\subsection{ Parts completed so far}
+
+We started by to generate a secret and public key depending on the size of the stream and the 
+security parameter. The next step was to generate the hash function according to each left and right leaf. 
+The hash function is computed as $ h(x,y) = Lx + Ry $ where \textbf{L, R} are selected randomly from $Z_{q}^{k \times m}$ and
+\textbf{x,y} are vectors of small norm.
+
+%genkey procedure generates a secret and public key.  The public key includes vectors \textbf{L} and \textbf{R} as well as $q$ and $\mathcal{U}$.
 
 %Questions on this step:
-%	\begin{enumerate}
-%	\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
-%	\end{enumerate}
+%\begin{enumerate}
+%\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
+%\end{enumerate}
 
 \subsection{ 2: Initialization}
-\label{initialize}
+%\label{initialize}
 
-	initialize takes the secret key and the public key generated in step 1, and an empty dictionary over the universe of elements and produces
+%initialize takes the secret key and the public key generated in step 1, and an empty dictionary over the universe of elements and produces
 
-	\begin{enumerate}
-	\item a table \textbf{T} with N entries (0 to N-1) to represent the Merkle tree leaves
-	\item initial state $d_{0} = \textbf{0} \in Z_{q}^{k}$ representing the root digest
-	\item a label for each internal node: $\lambda(v)$
-	\end{enumerate}
+%\begin{enumerate}
+%\item a table \textbf{T} with N entries (0 to N-1) to represent the Merkle tree leaves
+%\item initial state $d_{0} = \textbf{0} \in Z_{q}^{k}$ representing the root digest
+%\item a label for each internal node: $\lambda(v)$
+%\end{enumerate}
 
-%	Questions on this step:
-%	\begin{enumerate}
-%	\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
-%	\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
-%	\end{enumerate}
+%Questions on this step:
+%\begin{enumerate}
+%\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
+%\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
+%\end{enumerate}
 
 \subsection{ 3: updateVerifier}
-\subsubsection{Inputs}
-	\begin{enumerate}
-	\item $x$ where $x \in \mathcal{U}$ is the current element of the stream
-	\item $d_{h}$ current state of $auth(D_{h})$
-	\item sk and pk - secret and public key \\
-	\end{enumerate}
-
-\subsubsection{Outputs}
-	\begin{enumerate}
-	\item $d_{h + 1}$ updated state of $auth(D_{h + 1})$
-	\end{enumerate}
-
-	\begin{equation}
-		d_{h+1} = d_{h} + \mathcal{D}_{r}(x, \textbf{1})
-	\end{equation}
-
-\subsubsection{Example}
-
-	\begin{itemize}
-	\item $\mathcal{U} = \{ 0, 1, 2, 3, 4, 5, 6, 7\}$ similar to example in figure 7.1
-	\item Suppose $x = 3$
-	\end{itemize}
-
-	Calculating $d_{h + 1}$ reduces to calculating 	$ \mathcal{D}_{r}(x, \textbf{1})$ because $d_{h}$ is known.  \\
-
-	In our case, we'd like to calculate $\mathcal{D}_{r}(3, \textbf{1})$, i.e. the digest of the root node with respect 3. \\
-
-	By definition 10, 
-	\begin{equation}
-	\mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b}(\mathcal{D}_{v_{11}}(3, \textbf{x}_{3})) 
-	\end{equation}
-where $v_{11}$ is the left child of $r$, and $\mathcal{D}_{v_{11}}$ is the partial digest of $v_{11}$.
-
-	This procedure continues for each internal node eventually reaching the leaf node. The partial digest of a leaf node, in our case, that representing the element $x = 3$ is given by:
-
-	\begin{equation}
-	\mathcal{D}_{3}(3, \textbf{1}) = \textbf{1}
-	\end{equation}
-
-	Substituting the calculated partial digests for each node in the path from the root to the updated leaf node results in:
-
-	\begin{equation}
-	\mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
-	\end{equation}
-
-	Thus, the new digest for the entire tree is given by:
-	\begin{equation}
-	d_{h+1} = d_{h} + \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
-	\end{equation}
-
-%\subsection{Questions}
+%\subsubsection{Inputs}
 %\begin{enumerate}
-%\item What are the roles of the secret and public key as inputs?
+%\item $x$ where $x \in \mathcal{U}$ is the current element of the stream
+%\item $d_{h}$ current state of $auth(D_{h})$
+%\item sk and pk - secret and public key \\
+%\end{enumerate}
+% 
+% \subsubsection{Outputs}
+% \begin{enumerate}
+% \item $d_{h + 1}$ updated state of $auth(D_{h + 1})$
+% \end{enumerate}
+% 
+% \begin{equation}
+% d_{h+1} = d_{h} + \mathcal{D}_{r}(x, \textbf{1})
+% \end{equation}
+% 
+% \subsubsection{Example}
+% 
+% \begin{itemize}
+% \item $\mathcal{U} = \{ 0, 1, 2, 3, 4, 5, 6, 7\}$ similar to example in figure 7.1
+% \item Suppose $x = 3$
+% \end{itemize}
+% 
+% Calculating $d_{h + 1}$ reduces to calculating 	$ \mathcal{D}_{r}(x, \textbf{1})$ because $d_{h}$ is known.  \\
+% 
+% In our case, we'd like to calculate $\mathcal{D}_{r}(3, \textbf{1})$, i.e. the digest of the root node with respect 3. \\
+% 
+% By definition 10, 
+% \begin{equation}
+% \mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b}(\mathcal{D}_{v_{11}}(3, \textbf{x}_{3})) 
+% \end{equation}
+% where $v_{11}$ is the left child of $r$, and $\mathcal{D}_{v_{11}}$ is the partial digest of $v_{11}$.
+% 
+% This procedure continues for each internal node eventually reaching the leaf node. The partial digest of a leaf node, in our case, that representing the element $x = 3$ is given by:
+% 
+% \begin{equation}
+% \mathcal{D}_{3}(3, \textbf{1}) = \textbf{1}
+% \end{equation}
+% 
+% Substituting the calculated partial digests for each node in the path from the root to the updated leaf node results in:
+% 
+% \begin{equation}
+% \mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
+% \end{equation}
+% 
+% Thus, the new digest for the entire tree is given by:
+% \begin{equation}
+% d_{h+1} = d_{h} + \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
+% \end{equation}
+% 
+% %\subsection{Questions}
+% %\begin{enumerate}
+% %\item What are the roles of the secret and public key as inputs?
 In the previous email, you said the prover generates the secret key and does not output it.
 Therefore, the verifier can't have any value here for $sk$ besides null.
-%\end{enumerate}
-
-\subsection{ 4: updateProver}
-\subsubsection{Input}
-	\begin{itemize}
-	\item $x$: $x \in \mathcal{U}$ is the current element of the stream
-	\item $pk$:  public key
-	\item $D_{h}$: The table, \textbf{T} which represents leaves of the Merkle tree described by the scheme
-	\item $auth(D_{h})$: The auxiliary information allowing authentication of the elements in $D_{h}$ including all labels $\lambda(v_{i})$ for internals nodes and $d_{h}$, the root digest of the whole tree.
-	\end{itemize}
-\subsubsection{Output}
-
-	\begin{itemize}
-	\item $D_{h + 1}$: Updated table \textbf{T}
-	\item $auth(D_{h + 1})$: Updated labels and root digest
-	\end{itemize}
-
-\subsubsection{Example}
-
-Using the same scenario as the verifier example where $x = 3$, the updated labels is given by:
-
-\begin{equation}
-\lambda(v_{i}) = \lambda(v_{i}) + \textbf{b}( \mathcal{D}_{v_{i}}(3, \textbf{1}) ) \textrm{ for } i = \ell, \ell -1, \ldots, 1  
-\end{equation}
-
-Calculating each $\lambda(v_{i})$ is straight forward.  In the last example, we understood how to calculate a partial digest, and the $\textbf{b}()$ function is well-defined.
+% %\end{enumerate}
+% 
+% \subsection{ 4: updateProver}
+% \subsubsection{Input}
+% \begin{itemize}
+% \item $x$: $x \in \mathcal{U}$ is the current element of the stream
+% \item $pk$:  public key
+% \item $D_{h}$: The table, \textbf{T} which represents leaves of the Merkle tree described by the scheme
+% \item $auth(D_{h})$: The auxiliary information allowing authentication of the elements in $D_{h}$ including all labels $\lambda(v_{i})$ for internals nodes and $d_{h}$, the root digest of the whole tree.
+% \end{itemize}
+% \subsubsection{Output}
+% 
+% \begin{itemize}
+% \item $D_{h + 1}$: Updated table \textbf{T}
+% \item $auth(D_{h + 1})$: Updated labels and root digest
+% \end{itemize}
+% 
+% \subsubsection{Example}
+% 
+% Using the same scenario as the verifier example where $x = 3$, the updated labels is given by:
+% 
+% \begin{equation}
+% \lambda(v_{i}) = \lambda(v_{i}) + \textbf{b}( \mathcal{D}_{v_{i}}(3, \textbf{1}) ) \textrm{ for } i = \ell, \ell -1, \ldots, 1  
+% \end{equation}
+% 
+% Calculating each $\lambda(v_{i})$ is straight forward.  In the last example, we understood how to calculate a partial digest, and the $\textbf{b}()$ function is well-defined.
 
@@
 
-\section{Matrix Multiplication in Ruby}
-
-\subsection{The \texttt{[row $\times$ column]} multiplication:}
-	\begin{itemize}
-	\item $ C = A_{n \times n} \times B_{n \times n}$
-	\item where $ c_{ij} = \sum\limits_{k=1}^n a_{ik} . b_{kj}   $
-	\item complexity : $\Theta(n^3)$ \\
-	\end{itemize}
-
-\subsection{Divide and Conquer}
-
-	\begin{itemize}
-
-        \item Each of the matrices is divided into four $\frac{n}{2} \times \frac{n}{2}$
-	\item The multiplication of two $ n \times  n $ matrices is reduced to 8 multiplications of 
-	two $\frac{n}{2} \times \frac{n}{2}$ matrices and four $\frac{n}{2} \times \frac{n}{2}$ matrix additions.
-	\item recurrence : $T(n) = 8.T. \frac{n}{2} + \Theta(n^2) \rightarrow \Theta(n^3)$ \\
-	\end{itemize}
-
-\subsection{Strassen's Algorithm}
-
-	\begin{itemize}
-
-        \item Bulky for implementation and not numerically stable
-        \item Useful and advantageous for large matrices
-	\item $ \Theta(n^{log7}) \approx \Theta(n^{2.807})$ \\
-	\end{itemize}
-
-\subsection{Coppersmith Winograd Algorithm}
-
-	\begin{itemize}
-
-        \item Theoretical optimum; never used in practice.
-	\item $ \Theta(n^{2.376}) $ \\
-	\end{itemize}
+% \section{Matrix Multiplication in Ruby}
+% 
+% 
+% \subsection{The \texttt{[row $\times$ column]} multiplication:}
+% \begin{itemize}
+% \item $ C = A_{n \times n} \times B_{n \times n}$
+% \item where $ c_{ij} = \sum\limits_{k=1}^n a_{ik} . b_{kj}   $
+% \item complexity : $\Theta(n^3)$ \\
+% \end{itemize}
+% 
+% 
+% \subsection{Divide and Conquer}
+% 
+% \begin{itemize}
+% 
+% \item Each of the matrices is divided into four $\frac{n}{2} \times \frac{n}{2}$
+% \item The multiplication of two $ n \times  n $ matrices is reduced to 8 multiplications of 
+% two $\frac{n}{2} \times \frac{n}{2}$ matrices and four $\frac{n}{2} \times \frac{n}{2}$ matrix additions.
+% \item recurrence : $T(n) = 8.T. \frac{n}{2} + \Theta(n^2) \rightarrow \Theta(n^3)$ \\
+% \end{itemize}
+% 
+% \subsection{Strassen's Algorithm}
+% 
+% \begin{itemize}
+% 
+% \item Bulky for implementation and not numerically stable
+% \item Useful and advantageous for large matrices
+% \item $ \Theta(n^{log7}) \approx \Theta(n^{2.807})$ \\
+% \end{itemize}
+% 
+% \subsection{Coppersmith Winograd Algorithm}
+% 
+% \begin{itemize}
+% 
+% \item Theoretical optimum; never used in practice.
+% \item $ \Theta(n^{2.376}) $ \\
+% \end{itemize}
 

### 2012-11-02 doc/milestone_draft1.tex
milestone_draft

@@
 Taking into account that the bottomline of our scheme is the Merkle tree structure and using lattice based hash
-function with homomorphic property, we are implementing the algorithm in Ruby language at this stage and 
-defining classes for actually testing and configuring different steps in the algorithm. \\
+function with homomorphic property, we are implementing the algorithm in Python or Ruby language at this stage and 
+actually testing and configuring different steps in the algorithm i.e. \texttt{genkey, initialize, updateVerifier, updateProver, query, verify}.   \\
 
-Languages such as Python or Ruby support
+Languages such as Python or Ruby have support for their community of users that grows up and that
+is one of its most important strengths. Ruby has a vibrant and growing community that is friendly towards
+people of all skill levels. Such groups allow users to network with other Ruby
+programmers.  \\
 
-\subsection{ Parts completed so far}
+As for Python, NumPy is an extension module which provides efficient operations on arrays of data allowing
+Python to operate as a high level language for manipulating numerical data.
+In addition, SciPy is a set of open source scientific and numerical tools for Python. It can 
+support special functions, integration, differential equation solvers, etc. \\
 
-We started by to generate a secret and public key depending on the size of the stream and the 
+We are also thinking to use some software system such as EventMachine that is designed for writing
+scalable applications in Ruby. EventMachine has been used for making various libraries and frameworks
+in which concurrency is considered. Examples of such application could be
+web servers and proxies, email and IM, authentication and authorization processors and 
+network applications. \\
+
+In our work, we have generated a secret and public key depending on the size of the stream and the 
 security parameter. The next step was to generate the hash function according to each left and right leaf. 
 The hash function is computed as $ h(x,y) = Lx + Ry $ where \textbf{L, R} are selected randomly from $Z_{q}^{k \times m}$ and
-\textbf{x,y} are vectors of small norm.
+\textbf{x,y} are vectors of small norm. \\
 
@@
 
-\subsection{ 2: Initialization}
+%\subsection{ 2: Initialization}
 %\label{initialize}
@@
 
-\subsection{ 3: updateVerifier}
+%\subsection{ 3: updateVerifier}
 %\subsubsection{Inputs}

### 2012-11-04 doc/milestone_draft1.tex
One sentence per line

@@
 %In the following, we will represent the steps that the streaming authenticated data structure follows to implement range search queries.
-Taking into account that the bottomline of our scheme is the Merkle tree structure and using lattice based hash
-function with homomorphic property, we are implementing the algorithm in Python or Ruby language at this stage and 
-actually testing and configuring different steps in the algorithm i.e. \texttt{genkey, initialize, updateVerifier, updateProver, query, verify}.   \\
+Taking into account that the bottomline of our scheme is the Merkle tree structure and using lattice based hash function with homomorphic property, we are implementing the algorithm in ython or Ruby language at this stage and actually testing and configuring different steps in the algorithm i.e.
+\texttt{genkey, initialize, updateVerifier, updateProver, query, verify}.
 
 Languages such as Python or Ruby have support for their community of users that grows up and that
-is one of its most important strengths. Ruby has a vibrant and growing community that is friendly towards
-people of all skill levels. Such groups allow users to network with other Ruby
-programmers.  \\
+is one of its most important strengths.
+Ruby has a vibrant and growing community that is friendly towards people of all skill levels.
+Such groups allow users to network with other Ruby programmers.
 
-As for Python, NumPy is an extension module which provides efficient operations on arrays of data allowing
-Python to operate as a high level language for manipulating numerical data.
-In addition, SciPy is a set of open source scientific and numerical tools for Python. It can 
-support special functions, integration, differential equation solvers, etc. \\
+As for Python, NumPy is an extension module which provides efficient operations on arrays of data allowing Python to operate as a high level language for manipulating numerical data.
+In addition, SciPy is a set of open source scientific and numerical tools for Python.
+It can support special functions, integration, differential equation solvers, etc.
 
 We are also thinking to use some software system such as EventMachine that is designed for writing
-scalable applications in Ruby. EventMachine has been used for making various libraries and frameworks
-in which concurrency is considered. Examples of such application could be
-web servers and proxies, email and IM, authentication and authorization processors and 
-network applications. \\
+scalable applications in Ruby.
+EventMachine has been used for making various libraries and frameworks in which concurrency is considered.
+Examples of such application could be web servers and proxies, email and IM, authentication and authorization processors and network applications.
 
-In our work, we have generated a secret and public key depending on the size of the stream and the 
-security parameter. The next step was to generate the hash function according to each left and right leaf. 
+In our work, we have generated a secret and public key depending on the size of the stream and the security parameter.
+The next step was to generate the hash function according to each left and right leaf.
 The hash function is computed as $ h(x,y) = Lx + Ry $ where \textbf{L, R} are selected randomly from $Z_{q}^{k \times m}$ and
-\textbf{x,y} are vectors of small norm. \\
+\textbf{x,y} are vectors of small norm.
 
@@
 
-\section{Implementation Issues\\}
+\section{Implementation Issues}
 \subsection{Proof of Concept}
@@
 
-\subsection{Applicable Libraries\\}
+\subsection{Applicable Libraries}
 \begin{enumerate}
@@
 arbitrary precision floating point arithmetic. \texttt{http://shoup.net/ntl/}
-\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of
-the important concrete classes of 
-structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, 
-module theory and the theory of algebras. In addition, certain
-families of structures from algebraic 
-geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
+\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of the important concrete classes of structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, module theory and the theory of algebras.
+In addition, certain families of structures from algebraic geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
 \end{enumerate}
 
-\subsection{ Library Prototype\\}
+\subsection{ Library Prototype}
 
 We decided to start a feasible implementation of the model by trying to find the right library
-for matrix operations and arbitrary precision floating point arithmetics. In our investigations,
-we had the following observations :
+for matrix operations and arbitrary precision floating point arithmetics.
+In our investigations, we had the following observations :
 
@@
 \item MAGMA is a collection of linear algebra GPU accelerated libraries designed and implemented
-by the team that developed LAPACK. Since we were not completely sure about MAGMA's dependency on
-CUDA and other related libraries we did not go into further detail at this stage.
+by the team that developed LAPACK.
+Since we were not completely sure about MAGMA's dependency on CUDA and other related libraries we did not go into further detail at this stage.
 \item We decided to work with Ruby (or Python) for the first implementation since they both have a

### 2012-11-04 doc/milestone_draft1.tex
Added bibliography tag
Using enumerate package.  Adds param to \begin{enumerate} for style of list
Submitted proposal state

--- 2012-11-02 doc/milestone_draft1.tex
@@
 \documentclass[11pt, letterpaper, oneside]{article}
-\usepackage{enumitem}
-\usepackage{calc}
+\usepackage{enumerate}
+\usepackage{ calc }
 \usepackage{ amssymb }
@@
 
-\title{\textbf{ CMCS 818J Project Milestone Report } \\  \small{Christopher Imbriano \& Behzad Koosha}}
+\title{CMCS 818J Project Milestone Report}
+\author{Behzad Koosha, Christopher Imbriano}
+
 \maketitle
@@
 
-As it has been described in the project proposal, the aim for this project is to present and implement a new model for verifying data structure queries in a streaming scenario.
-The proposed scheme gives rise to new cryptographic constructions with improved properties such as dynamic cryptographic accumulator; parallel online memory checking and a space efficient three party authenticated data structure.
+	A scheme for achieving a streaming authenticated data structure (SADS) was proposed by Papamanthou and Shi \cite{sads}.
+	The goal of this project is twofold.
+	First, we will implement an example of this data structure and second, we will build an application using the SADS primitive.
 
@@
 	The introduced scheme, streaming authenticated data structure, does not require any interaction between the client and the server while the stream is observed.
-Three important properties is obtained using this scheme : a.
-Independence of prover and verifier : the prover and the verifier update their status independently and with no interaction.
-b.
-Efficiency : It can be shown that an exponential improvement for many commonly queries in the prover's running time can be achieved.
-c.
-Expressiveness : a wide range of queries such as membership/ non-membership, successor, range search and frequencies can be supported using this scheme.
-
-\section{Previous Work}
-
-Most of the existing streaming verifiable protocols are efficient in terms of verifier complexity; it can be observed that in all of them the prover complexity is linear in the size of stream which this limits the applicability of the protocols.
-Other works include verifiable computation and authenticated data structures, which are not directly applicable to the streaming data set or their application results in increased complexities or interactive protocols.
+	Three important properties are obtained using this scheme :
+	\begin{enumerate}[a.]
+	\item Independence of prover and verifier : the prover and the verifier update their status independently and with no interaction.
+	\item Efficiency : It can be shown that an exponential improvement for many commonly queries in the prover's running time can be achieved.
+	\item Expressiveness : a wide range of queries such as membership/ non-membership, successor, range search and frequencies can be supported using this scheme.
+	\end{enumerate}
 
@@
 
-%In the following, we will represent the steps that the streaming authenticated data structure follows to implement range search queries.
-Taking into account that the bottomline of our scheme is the Merkle tree structure and using lattice based hash
-function with homomorphic property, we are implementing the algorithm in Ruby language at this stage and 
-defining classes for actually testing and configuring different steps in the algorithm. \\
+	The description of the SADS scheme reads much like a protocol, focusing on the behaviour of each of the prover and verifier during a streaming phase and a query phase.
+	Before tackling the problem at the protocol level however, we must first implement the basic data structures each of the verifier and prover will need to perform their roles.
+	This will be our starting point.
 
-Languages such as Python or Ruby support
+	The road map for this project is as follows:
+	\begin{enumerate}
+	\item Implement underlying Merkle tree with homomorphic hash function
+	\item Implement the SADS scheme
+	\item Develop proof-of-concept application using the SADS scheme implementation
+	\end{enumerate}
 
-\subsection{ Parts completed so far}
+	\subsection{Underlying Data Structures}
 
-We started by to generate a secret and public key depending on the size of the stream and the 
-security parameter. The next step was to generate the hash function according to each left and right leaf. 
-The hash function is computed as $ h(x,y) = Lx + Ry $ where \textbf{L, R} are selected randomly from $Z_{q}^{k \times m}$ and
-\textbf{x,y} are vectors of small norm.
+	A crucial component of the SADS algorithm relies on matrix multiplication, a relatively expensive computational operation.
+	When formulating our plan of action for this project, we were recommended a few linear algebra libraries, which will be discussed in a later section.
+	While we recognize the importance of fast matrix operations and parallelism for production quality code, we thought that the added complexity, as well as unknowns with respect to implementation details, would be prohibitive.
+	Instead, we've decided to use an interpreted language and the simple matrix operations that come with the standard libraries to start.
+
+	Our language of choice for this first stage of the project is Ruby. Ruby has a number of advantages including:
+
+	\begin{enumerate}
+	\item Built-in support of matrix and vector operations
+	\item Vibrant developer community that supports both new and experienced developers through learning resources, tool development, etc...
+	\item Ability to write compiled extensions in C
+	\item Numeric and visualization library, SciRuby
+	\end{enumerate}
+
+	\subsection{SADS Scheme}
+
+	Once the underlying data structure is complete, the SADS scheme will use it as a primitive.
+	Though one of the scheme's strengths is the independence of the prover and verifier during the streaming setting, the two parties will necessarily need to communicate during the query phase.
+	Additionally, there may be communication between the two parties to initiate the streaming setting.
+	For these reasons, we anticipate creating an application which will require network communication as it will likely be the case that the verifier and prover are not collocated.
+
+	We have a few frameworks in mind for this task.
+	Two of our options at the moment are EventMachine and Node.js.
+	Node.js is a "platform for building fast, scalable network applications" and has been gaining popularity in the Ruby community. http://nodejs.org/
+	EventMachine is a "fast, simple event-processing library for Ruby programs."http://rubyeventmachine.com/
+
+% In our work, we have generated a secret and public key depending on the size of the stream and the security parameter.
+% The next step was to generate the hash function according to each left and right leaf.
+% The hash function is computed as $ h(x,y) = Lx + Ry $ where \textbf{L, R} are selected randomly from $Z_{q}^{k \times m}$ and
+% \textbf{x,y} are vectors of small norm.
 
@@
 
-%Questions on this step:
-%\begin{enumerate}
-%\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
-%\end{enumerate}
+\section{Looking Forward}
 
-\subsection{ 2: Initialization}
-%\label{initialize}
+	\subsection{Production Quality Libraries}
+	It is yet to be seen whether the existing Ruby matrix libraries, either the built in or SciRuby, will be sufficiently efficient for the needs of the SADS scheme.
+	That's not to say the current choices of technologies will not produce production quality code but, as mentioned earlier, we have been investigating some optimized libraries listed here.
 
-%initialize takes the secret key and the public key generated in step 1, and an empty dictionary over the universe of elements and produces
-
-%\begin{enumerate}
-%\item a table \textbf{T} with N entries (0 to N-1) to represent the Merkle tree leaves
-%\item initial state $d_{0} = \textbf{0} \in Z_{q}^{k}$ representing the root digest
-%\item a label for each internal node: $\lambda(v)$
-%\end{enumerate}
-
-%Questions on this step:
-%\begin{enumerate}
-%\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
-%\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
-%\end{enumerate}
-
-\subsection{ 3: updateVerifier}
-%\subsubsection{Inputs}
-%\begin{enumerate}
-%\item $x$ where $x \in \mathcal{U}$ is the current element of the stream
-%\item $d_{h}$ current state of $auth(D_{h})$
-%\item sk and pk - secret and public key \\
-%\end{enumerate}
-% 
-% \subsubsection{Outputs}
-% \begin{enumerate}
-% \item $d_{h + 1}$ updated state of $auth(D_{h + 1})$
-% \end{enumerate}
-% 
-% \begin{equation}
-% d_{h+1} = d_{h} + \mathcal{D}_{r}(x, \textbf{1})
-% \end{equation}
-% 
-% \subsubsection{Example}
-% 
-% \begin{itemize}
-% \item $\mathcal{U} = \{ 0, 1, 2, 3, 4, 5, 6, 7\}$ similar to example in figure 7.1
-% \item Suppose $x = 3$
-% \end{itemize}
-% 
-% Calculating $d_{h + 1}$ reduces to calculating 	$ \mathcal{D}_{r}(x, \textbf{1})$ because $d_{h}$ is known.  \\
-% 
-% In our case, we'd like to calculate $\mathcal{D}_{r}(3, \textbf{1})$, i.e. the digest of the root node with respect 3. \\
-% 
-% By definition 10, 
-% \begin{equation}
-% \mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b}(\mathcal{D}_{v_{11}}(3, \textbf{x}_{3})) 
-% \end{equation}
-% where $v_{11}$ is the left child of $r$, and $\mathcal{D}_{v_{11}}$ is the partial digest of $v_{11}$.
-% 
-% This procedure continues for each internal node eventually reaching the leaf node. The partial digest of a leaf node, in our case, that representing the element $x = 3$ is given by:
-% 
-% \begin{equation}
-% \mathcal{D}_{3}(3, \textbf{1}) = \textbf{1}
-% \end{equation}
-% 
-% Substituting the calculated partial digests for each node in the path from the root to the updated leaf node results in:
-% 
-% \begin{equation}
-% \mathcal{D}_{r}(3, \textbf{1}) = \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
-% \end{equation}
-% 
-% Thus, the new digest for the entire tree is given by:
-% \begin{equation}
-% d_{h+1} = d_{h} + \textbf{L} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{R} \cdot \textbf{b} ( \textbf{1} ) )  )
-% \end{equation}
-% 
-% %\subsection{Questions}
-% %\begin{enumerate}
-% %\item What are the roles of the secret and public key as inputs?
-In the previous email, you said the prover generates the secret key and does not output it.
-Therefore, the verifier can't have any value here for $sk$ besides null.
-% %\end{enumerate}
-% 
-% \subsection{ 4: updateProver}
-% \subsubsection{Input}
-% \begin{itemize}
-% \item $x$: $x \in \mathcal{U}$ is the current element of the stream
-% \item $pk$:  public key
-% \item $D_{h}$: The table, \textbf{T} which represents leaves of the Merkle tree described by the scheme
-% \item $auth(D_{h})$: The auxiliary information allowing authentication of the elements in $D_{h}$ including all labels $\lambda(v_{i})$ for internals nodes and $d_{h}$, the root digest of the whole tree.
-% \end{itemize}
-% \subsubsection{Output}
-% 
-% \begin{itemize}
-% \item $D_{h + 1}$: Updated table \textbf{T}
-% \item $auth(D_{h + 1})$: Updated labels and root digest
-% \end{itemize}
-% 
-% \subsubsection{Example}
-% 
-% Using the same scenario as the verifier example where $x = 3$, the updated labels is given by:
-% 
-% \begin{equation}
-% \lambda(v_{i}) = \lambda(v_{i}) + \textbf{b}( \mathcal{D}_{v_{i}}(3, \textbf{1}) ) \textrm{ for } i = \ell, \ell -1, \ldots, 1  
-% \end{equation}
-% 
-% Calculating each $\lambda(v_{i})$ is straight forward.  In the last example, we understood how to calculate a partial digest, and the $\textbf{b}()$ function is well-defined.
-
-%\subsection{Questions}
-
-%From the algorithm $updateProver$ on page 10,
-
-%\begin{quote}
-%The algorithm sets $\textbf{T}[x]$ = $\textbf{T}[x] + \textbf{1}$, outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$. Let now
-%$v_{\ell}$, $\ldots$, $v_{1}$ be the path in the lattice-based Merkle tree $T$ from node $v_{\ell}$ ($v_{\ell}$ stores the value  $\textbf{T}[x]$) to the child $v_{1}$ of the root $r$ of $T$.
-%\end{quote}
-
-%\begin{enumerate}
-%\item How is  $\ell$ calculated in our scenario?
-
-%\item What is meant by ``$v_{\ell}$ stores the value $\textbf{T}[x]$''?
-Is $v_{\ell}$ the leaf node of the Merkle tree?
-If so, equation 5.5 would update a label for $v_{\ell}$, but that contradicts our understanding of the $initialize$ algorithm which assigns labels for only internal nodes.
-
-%\item Looking at the tree in Figure 7.1, we think the path is ($v_{22}$, $v_{11}$) where $\ell = 2$, because these are the only internal nodes for which there is a label to update.
-(The root has a digest, $d_{h}$, and the leaf node has an entry in $\textbf{T}$.)  Additionally, the algorithm states the path is from $v_{\ell}$ to ``the child $v_{1}$ of the root $r$ of $T$" suggesting that the last element of the path in question is the child of the root node, not the root node itself.
-Is this correct?
-
-%\item A notation question about the statement ``outputting the updated $\textbf{T}[x]$ as $D_{h + 1}$".
-Isn't \textbf{T} the whole table, whereas $\textbf{T}[x]$ is an element in \textbf{T}.
-If so, is there not a type mismatch in that statement, where $D_{h + 1}$ is the whole data structure and $\textbf{T}[x]$ is just one element?
-
-%\item When/where does the root digest get updated (especially if our understanding of the path in the questions 2 and 3 is correct, i.e.
-the root is not included in that path)?
-Shouldn't there be a similar update for the root or the root digest?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Implementation Issues\\}
-\subsection{Proof of Concept}
 	\begin{enumerate}
-\item In order to implement a streaming authenticated data structure based on lattices, a fast
-arithmetic library for matrices and vector is required.
-\item Matrix multiplication mod \texttt{q} ( \texttt{q poly(k)}) is needed.
+	\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length integers, for vectors, matrices and polynomials over the integers and over the finite fields and for arbitrary precision floating point arithmetic.
+\texttt{http://shoup.net/ntl/}
+	\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of the important concrete classes of structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, module theory and the theory of algebras.
+	In addition, certain families of structures from algebraic geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
 	\end{enumerate}
 
-\subsection{Applicable Libraries\\}
-\begin{enumerate}
-\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length
-integers, for vectors, matrices and polynomials over the integers and over the finite fields and for 
-arbitrary precision floating point arithmetic. \texttt{http://shoup.net/ntl/}
-\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of
-the important concrete classes of 
-structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, 
-module theory and the theory of algebras. In addition, certain
-families of structures from algebraic 
-geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
-\end{enumerate}
-
-\subsection{ Library Prototype\\}
-
-We decided to start a feasible implementation of the model by trying to find the right library
-for matrix operations and arbitrary precision floating point arithmetics. In our investigations,
-we had the following observations :
-
-\begin{itemize}
-
-\item It seems that there has been more recent updates and user discussion links regarding MAGMA
-and the library has attracted more draw in terms of being more up to dated and user-friendly.
-\item MAGMA is a collection of linear algebra GPU accelerated libraries designed and implemented
-by the team that developed LAPACK. Since we were not completely sure about MAGMA's dependency on
-CUDA and other related libraries we did not go into further detail at this stage.
-\item We decided to work with Ruby (or Python) for the first implementation since they both have a
-standard library documentation for matrices.
-\item Starting with this idea, we can smoothly find the best matching library for our system 
-model which helps us to know exactly what requirements we need and which library matches these 
-requirements accordingly.
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\end{enumerate}
-
-% \section{Matrix Multiplication in Ruby}
-% 
-% 
-% \subsection{The \texttt{[row $\times$ column]} multiplication:}
-% \begin{itemize}
-% \item $ C = A_{n \times n} \times B_{n \times n}$
-% \item where $ c_{ij} = \sum\limits_{k=1}^n a_{ik} . b_{kj}   $
-% \item complexity : $\Theta(n^3)$ \\
-% \end{itemize}
-% 
-% 
-% \subsection{Divide and Conquer}
-% 
-% \begin{itemize}
-% 
-% \item Each of the matrices is divided into four $\frac{n}{2} \times \frac{n}{2}$
-% \item The multiplication of two $ n \times  n $ matrices is reduced to 8 multiplications of 
-% two $\frac{n}{2} \times \frac{n}{2}$ matrices and four $\frac{n}{2} \times \frac{n}{2}$ matrix additions.
-% \item recurrence : $T(n) = 8.T. \frac{n}{2} + \Theta(n^2) \rightarrow \Theta(n^3)$ \\
-% \end{itemize}
-% 
-% \subsection{Strassen's Algorithm}
-% 
-% \begin{itemize}
-% 
-% \item Bulky for implementation and not numerically stable
-% \item Useful and advantageous for large matrices
-% \item $ \Theta(n^{log7}) \approx \Theta(n^{2.807})$ \\
-% \end{itemize}
-% 
-% \subsection{Coppersmith Winograd Algorithm}
-% 
-% \begin{itemize}
-% 
-% \item Theoretical optimum; never used in practice.
-% \item $ \Theta(n^{2.376}) $ \\
-% \end{itemize}
-
-\section{Future Work}
-
-For the remaining time in the semester, we plan to implement our idea in Ruby and then move on to implenting the structure to the right library such as MAGMA.
-We aim to design and implement the scheme for applications that can fully enable efficient verification of data structure queries on a stream.
-In addition, we will be meeting with Elaine (and Babis via Skype) to discuss and exchange thoughts about our approach to solve the problem as we have already practiced it before.
+	\subsection{Semester Outlook}
+		While the ultimate goal of the project is to produce a proof-of-concept application which utilizes a SADS scheme, it is best to set manage expectations.
+		Our baseline target for this semester is to complete the first two layers described above, that is the underlying data structure and the SADS scheme primitive.
+		Having completed these two components, we can then establish time line during which the proof-of-concept application can be hashed out and developed.
+		We hope to leverage the Ruby and broader open-source community by packaging this application, or just the SADS scheme primitive, into a library for others to use and provide feedback.
 
@@
 
+\begin{thebibliography}{9}
+	\bibitem{sads} Elaine Shi, Charalampos Papamanthou \emph{Streaming Authenticated Data Structures}, Date Unknown
+\end{thebibliography}
+
 \end{document}

### 2012-11-04 doc/algorithm.tex
Adding commented out stuff from the milestone

--- 2012-10-02 doc/scheme example.tex
@@
 \documentclass[11pt, letterpaper, oneside]{article}
-\usepackage{enumitem}
+\usepackage{enumerate}
 \usepackage{calc}
@@
 \begin{document}
-\section{Algortihm 1: Generate Keys}
 
-	genkey procedure generates a secret and public key.  The public key includes vectors \textbf{L} and \textbf{R} as well as $q$ and $\mathcal{U}$.
+Questions on this step:
+\begin{enumerate}
+\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
+\end{enumerate}
 
-%Questions on this step:
-%	\begin{enumerate}
-%	\item	 In the first algorithm defined on page 10, what does the character in 'sk =  \O' represent? Is that the null set?
-%	\end{enumerate}
-
-\section{Algorithm 2: Initialization}
+\subsection{ 2: Initialization}
 \label{initialize}
@@
 
-%	Questions on this step:
-%	\begin{enumerate}
-%	\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
-%	\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
-%	\end{enumerate}
+Questions on this step:
+\begin{enumerate}
+\item What are the dimensions of table \textbf{T}?  ( $N \times ?$ )
+\item What does q mean in the context of $Z_{q}^{k}$?  Is this a set of $k \times 1$ vectors or a set of $q \times k$ matrices?
+\end{enumerate}
 
-\section{Algorithm 3: updateVerifier}
-\subsection{Inputs}
+\subsection{ 3: updateVerifier}
+\subsubsection{Inputs}
 \begin{enumerate}
@@
 
-\subsection{Outputs}
+\subsubsection{Outputs}
 \begin{enumerate}
@@
 
-\subsection{Questions}
-\begin{enumerate}
-\item What are the roles of the secret and public key as inputs?
+%\subsection{Questions}
+%\begin{enumerate}
+%\item What are the roles of the secret and public key as inputs?
 In the previous email, you said the prover generates the secret key and does not output it.
 Therefore, the verifier can't have any value here for $sk$ besides null.
-\end{enumerate}
+%\end{enumerate}
 
-\section{Algorithm 4: updateProver}
-\subsection{Input}
+\subsection{ 4: updateProver}
+\subsubsection{Input}
 \begin{itemize}
@@
 \end{itemize}
-\subsection{Output}
+\subsubsection{Output}
 
@@
 
+\section{Matrix Multiplication in Ruby}
+
+\subsection{The \texttt{[row $\times$ column]} multiplication:}
+\begin{itemize}
+\item $ C = A_{n \times n} \times B_{n \times n}$
+\item where $ c_{ij} = \sum\limits_{k=1}^n a_{ik} . b_{kj}   $
+\item complexity : $\Theta(n^3)$ \\
+\end{itemize}
+
+\subsection{Divide and Conquer}
+
+\begin{itemize}
+
+\item Each of the matrices is divided into four $\frac{n}{2} \times \frac{n}{2}$
+\item The multiplication of two $ n \times  n $ matrices is reduced to 8 multiplications of
+two $\frac{n}{2} \times \frac{n}{2}$ matrices and four $\frac{n}{2} \times \frac{n}{2}$ matrix additions.
+\item recurrence : $T(n) = 8.T. \frac{n}{2} + \Theta(n^2) \rightarrow \Theta(n^3)$ \\
+\end{itemize}
+
+\subsection{Strassen's Algorithm}
+
+\begin{itemize}
+
+\item Bulky for implementation and not numerically stable
+\item Useful and advantageous for large matrices
+\item $ \Theta(n^{log7}) \approx \Theta(n^{2.807})$ \\
+\end{itemize}
+
+\subsection{Coppersmith Winograd Algorithm}
+
+\begin{itemize}
+
+\item Theoretical optimum; never used in practice.
+\item $ \Theta(n^{2.376}) $ \\
+\end{itemize}
+
 \end{document}

### 2012-11-04 test/helper.rb
Using minitest instead of testunit

@@ -1,6 +1,4 @@
 require 'simplecov'
-#SimpleCov.start if ENV['COVERAGE']
-
 require 'rubygems'
 require 'bundler'
 begin
@@ -10,12 +8,18 @@ rescue Bundler::BundlerError => e
   $stderr.puts "Run `bundle install` to install missing gems"
   exit e.status_code
 end
-require 'test/unit'
-require 'shoulda'
+
+require 'minitest/unit'
+require "minitest/reporters"
+MiniTest::Reporters.use! MiniTest::Reporters::SpecReporter.new
+
+#require 'shoulda'
 
 $LOAD_PATH.unshift(File.join(File.dirname(__FILE__), '..', 'lib'))
 $LOAD_PATH.unshift(File.dirname(__FILE__))
 require 'sads'
 
-class Test::Unit::TestCase
+class MiniTest::Unit::TestCase
 end
+
+MiniTest::Unit.autorun

### 2012-11-04 test/test_sads.rb

@@ -1,6 +1,6 @@
 require 'helper'
 
-class TestSads < Test::Unit::TestCase
+class TestSads < MiniTest::Unit::TestCase
 	def setup
 		@sad = Sads.new
 	end

### 2012-11-04 test/helper.rb
Using minimal gems for now until we figure out proper use of groups in Gemfile

@@ -10,8 +10,8 @@ rescue Bundler::BundlerError => e
 end
 
 require 'minitest/unit'
-require "minitest/reporters"
-MiniTest::Reporters.use! MiniTest::Reporters::SpecReporter.new
+# require "minitest/reporters"
+# MiniTest::Reporters.use! MiniTest::Reporters::SpecReporter.new
 
 #require 'shoulda'
 

### 2012-11-05 test/helper.rb
Adding minitest reporters gem for coloring output

@@ -10,8 +10,8 @@ rescue Bundler::BundlerError => e
 end
 
 require 'minitest/unit'
-# require "minitest/reporters"
-# MiniTest::Reporters.use! MiniTest::Reporters::SpecReporter.new
+require "minitest/reporters"
+MiniTest::Reporters.use! MiniTest::Reporters::SpecReporter.new
 
 #require 'shoulda'
 

### 2012-11-05 test/helper.rb
Using DefaultReporter for minitest colorization. Skipping unimplemented tests instead of failing.

@@ -11,7 +11,10 @@ end
 
 require 'minitest/unit'
 require "minitest/reporters"
-MiniTest::Reporters.use! MiniTest::Reporters::SpecReporter.new
+require 'minitest/autorun'
+MiniTest::Reporters.use! MiniTest::Reporters::DefaultReporter.new
+
+#MiniTest::Reporters::SpecReporter.new
 
 #require 'shoulda'
 

### 2012-11-05 test/test_sads.rb

@@ -6,23 +6,26 @@ class TestSads < MiniTest::Unit::TestCase
 	end
 
 	def test_constructor
-		assert_not_nil(@sad)
+		@sad.wont_be_nil
 	end
 
 	# Leave test here: failing test at the moment
 	def test_calc_mu_q_0
-		assert(false, "Not implemented")
-		assert_equal(0, @sad.calculate_mu(0,0))
+		skip
+		# assert(false, "Not implemented")
+		# assert_equal(0, @sad.calculate_mu(0,0))
 	end
 
 	def test_mu_is_not_a_float
-		assert(false, "Not implemented")
+		skip
+		# assert(false, "Not implemented")
 	end
-	
+
 	def test_calc_q
-		assert(false, "Not implemented")
-	end 
-	
+		skip
+		# assert(false, "Not implemented")
+	end
+
 
 	def test_calc_mu
 		assert_equal(12, @sad.calculate_mu(2,8))
@@ -62,13 +65,15 @@ class TestSads < MiniTest::Unit::TestCase
 	end
 
 	def test_hash_elements_are_mod_q
-		@sad.hash(x,y).each do |item|
-			assert(item < @sad.q)
-		end
+		skip
+		# @sad.hash(x,y).each do |item|
+		# 	assert(item < @sad.q)
+		# end
 	end
 
 	def test_hash_result_is_a_matrix
-		assert_instance_of(Matrix, @sad.hash(x,y))
+		skip
+		# assert_instance_of(Matrix, @sad.hash(x,y))
 	end
 
 end

### 2012-11-05 test/helper.rb
Wrote test for mu is integer and testing basic properties of hash

@@ -21,6 +21,7 @@ MiniTest::Reporters.use! MiniTest::Reporters::DefaultReporter.new
 $LOAD_PATH.unshift(File.join(File.dirname(__FILE__), '..', 'lib'))
 $LOAD_PATH.unshift(File.dirname(__FILE__))
 require 'sads'
+require 'mathn'
 
 class MiniTest::Unit::TestCase
 end

### 2012-11-05 test/test_sads.rb

@@ -9,16 +9,16 @@ class TestSads < MiniTest::Unit::TestCase
 		@sad.wont_be_nil
 	end
 
-	# Leave test here: failing test at the moment
-	def test_calc_mu_q_0
-		skip
-		# assert(false, "Not implemented")
-		# assert_equal(0, @sad.calculate_mu(0,0))
-	end
-
 	def test_mu_is_not_a_float
-		skip
-		# assert(false, "Not implemented")
+		primes = Prime.instance
+
+		primes.take(10).each do |q|
+
+			(1..10).each do |k|
+
+				@sad.calculate_mu(k, q).must_be_instance_of Fixnum
+			end
+		end
 	end
 
 	def test_calc_q
@@ -64,16 +64,31 @@ class TestSads < MiniTest::Unit::TestCase
 		assert( false == @sad.L.eql?(@sad.R))
 	end
 
-	def test_hash_elements_are_mod_q
-		skip
-		# @sad.hash(x,y).each do |item|
-		# 	assert(item < @sad.q)
-		# end
+	def test_hash_result_is_a_matrix
+
+		# x and y have to be m entries "tall" if L & R are k x m
+		# m = mu / 2
+
+		mu = @sad.mu / 2
+
+		x = Matrix.column_vector( (0...mu).to_a )
+		y = Matrix.column_vector( (0...mu).to_a.map { |e| e * 2  } )
+
+		assert_instance_of(Matrix, @sad.hash(x,y))
 	end
 
-	def test_hash_result_is_a_matrix
-		skip
-		# assert_instance_of(Matrix, @sad.hash(x,y))
+	def test_hash_elements_are_mod_q
+
+		mu = @sad.mu / 2
+
+		x = Matrix.column_vector( (0...mu).to_a )
+		y = Matrix.column_vector( (0...mu).to_a.map { |e| e * 2  } )
+
+		@sad.hash(x,y).each do |item|
+			assert(item < @sad.q, "#{item} is >= #{@sad.q}")
+		end
 	end
+
+
 
 end

### 2012-11-05 lib/sads.rb
Whitespace

@@ -2,7 +2,7 @@
 require 'matrix'
 
 class Sads
-	
+
 	attr_accessor :labels, :leaves, :digest
 
 	# Left and Right vectors for algebraic hash function
@@ -33,11 +33,11 @@ class Sads
 	def initialize()
 		# Input parameteres?
 		#  # Size of Universe
-				
+
 		@k = 5
 		@q = 17
 		@mu   = calculate_mu(@k, @q)
-		
+
 		init_L_R
 
 	end
@@ -59,9 +59,9 @@ class Sads
 
 	def calculate_q(n, k)
 		# q is the smallest prime satisfying
-		# 
+		#
 		# q / log (q + 1) >= n * 2k * w( sqrt(k * log k))
-		
+
 		# Temporarily Set to arbirary value
 		n * k
 	end

### 2012-11-05 test/helper.rb

@@ -1,6 +1,7 @@
 require 'simplecov'
 require 'rubygems'
 require 'bundler'
+
 begin
   Bundler.setup(:default, :development)
 rescue Bundler::BundlerError => e

### 2012-11-06 test/sads_helper.rb
For helper methods for testing

@@ -1,0 +1,8 @@
+
+def LR_matrix(k, m, q)
+	Matrix.build(k, m) { rand q}
+end
+
+def node_label(m)
+	Matrix.column_vector( (0...m).to_a )
+end

### 2012-11-06 test/test_sads.rb
Using help method to generate test values for hash input parameters

@@ -1,4 +1,5 @@
 require 'helper'
+require 'sads_helper'
 
 class TestSads < MiniTest::Unit::TestCase
 	def setup
@@ -65,24 +66,19 @@ class TestSads < MiniTest::Unit::TestCase
 	end
 
 	def test_hash_result_is_a_matrix
+		m = @sad.mu / 2
 
-		# x and y have to be m entries "tall" if L & R are k x m
-		# m = mu / 2
-
-		mu = @sad.mu / 2
-
-		x = Matrix.column_vector( (0...mu).to_a )
-		y = Matrix.column_vector( (0...mu).to_a.map { |e| e * 2  } )
+		x = node_label(m)
+		y = node_label(m)
 
 		assert_instance_of(Matrix, @sad.hash(x,y))
 	end
 
 	def test_hash_elements_are_mod_q
+		m = @sad.mu / 2
 
-		mu = @sad.mu / 2
-
-		x = Matrix.column_vector( (0...mu).to_a )
-		y = Matrix.column_vector( (0...mu).to_a.map { |e| e * 2  } )
+		x = node_label(m)
+		y = node_label(m)
 
 		@sad.hash(x,y).each do |item|
 			assert(item < @sad.q, "#{item} is >= #{@sad.q}")

### 2012-11-06 test/helper.rb
Simplecov affected by bug in mathn

@@ -15,14 +15,13 @@ require "minitest/reporters"
 require 'minitest/autorun'
 MiniTest::Reporters.use! MiniTest::Reporters::DefaultReporter.new
 
-#MiniTest::Reporters::SpecReporter.new
+MiniTest::Reporters::SpecReporter.new
 
 #require 'shoulda'
 
 $LOAD_PATH.unshift(File.join(File.dirname(__FILE__), '..', 'lib'))
 $LOAD_PATH.unshift(File.dirname(__FILE__))
 require 'sads'
-require 'mathn'
 
 class MiniTest::Unit::TestCase
 end

### 2012-11-06 test/test_sads.rb

@@ -1,5 +1,6 @@
 require 'helper'
 require 'sads_helper'
+# require 'mathn'
 
 class TestSads < MiniTest::Unit::TestCase
 	def setup
@@ -11,7 +12,9 @@ class TestSads < MiniTest::Unit::TestCase
 	end
 
 	def test_mu_is_not_a_float
-		primes = Prime.instance
+		# primes = Prime.instance
+
+		primes = [2,3,5,7,11,13,17,19,23,29]
 
 		primes.take(10).each do |q|
 

### 2012-11-06 test/test_sads.rb
require prime instead of mathn'

@@ -1,6 +1,6 @@
 require 'helper'
 require 'sads_helper'
-# require 'mathn'
+require 'prime'
 
 class TestSads < MiniTest::Unit::TestCase
 	def setup
@@ -13,10 +13,9 @@ class TestSads < MiniTest::Unit::TestCase
 
 	def test_mu_is_not_a_float
 		# primes = Prime.instance
+		# primes = [2,3,5,7,11,13,17,19,23,29]
 
-		primes = [2,3,5,7,11,13,17,19,23,29]
-
-		primes.take(10).each do |q|
+		Prime.each (100) do |q|
 
 			(1..10).each do |k|
 

### 2012-11-06 lib/sads.rb
Sads initialize takes k and n parameters

@@ -30,12 +30,14 @@ class Sads
 	#beta
 	attr_reader :beta
 
-	def initialize()
+	def initialize(k, n)
 		# Input parameteres?
 		#  # Size of Universe
 
-		@k = 5
-		@q = 17
+		@k = k
+		@stream_bound_n = n
+
+		@q = calculate_q(k, n)
 		@mu   = calculate_mu(@k, @q)
 
 		init_L_R

### 2012-11-06 lib/sads.rb
Added binary_vector method

@@ -58,14 +58,45 @@ class Sads
 	end
 
 	# private below here
+	#
 
-	def calculate_q(n, k)
+
+
+
+	def binary_vector(x)
+
+		b_parts = Array.new
+
+		x.each do |ele|
+
+			# Make each element in x into binary form
+			# 	Add each bit to output vector
+
+			bin = ele.to_s(2)
+
+			(Math.log2(q).ceil - bin.length).times do
+				b_parts << 0
+			end
+
+			bin.each_char do |bit|
+				b_parts << bit.to_i
+
+			end
+		end
+
+		Matrix.column_vector(b_parts)
+	end
+
+	def calculate_q(k, n)
 		# q is the smallest prime satisfying
-		#
 		# q / log (q + 1) >= n * 2k * w( sqrt(k * log k))
+		#
+		# Using THETA notation, epsilon = 1
 
-		# Temporarily Set to arbirary value
-		n * k
+
+		# Using ceiling for now, until we actually find the
+		# smallest prime bigger than this quantity
+		(n * k * Math.log2(k) * Math.sqrt( k * Math.log2(k) )).ceil
 	end
 
 	def calculate_mu(k, q)

### 2012-11-06 test/sads_helper.rb
Renamed node_label to comlumn_vector, tested calc_q and initailized Sads with params

@@ -3,6 +3,6 @@ def LR_matrix(k, m, q)
 	Matrix.build(k, m) { rand q}
 end
 
-def node_label(m)
-	Matrix.column_vector( (0...m).to_a )
+def column_vector(size)
+	Matrix.column_vector( (0...size).to_a )
 end

### 2012-11-06 test/test_sads.rb

@@ -3,8 +3,9 @@ require 'sads_helper'
 require 'prime'
 
 class TestSads < MiniTest::Unit::TestCase
+
 	def setup
-		@sad = Sads.new
+		@sad = Sads.new(2,10)
 	end
 
 	def test_constructor
@@ -25,8 +26,9 @@ class TestSads < MiniTest::Unit::TestCase
 	end
 
 	def test_calc_q
-		skip
-		# assert(false, "Not implemented")
+
+		# assert_true( calculate_q(1024, 256)
+		assert_equal(29, @sad.calculate_q(2,10))
 	end
 
 
@@ -70,23 +72,40 @@ class TestSads < MiniTest::Unit::TestCase
 	def test_hash_result_is_a_matrix
 		m = @sad.mu / 2
 
-		x = node_label(m)
-		y = node_label(m)
+		x = column_vector(m)
+		y = column_vector(m)
 
 		assert_instance_of(Matrix, @sad.hash(x,y))
 	end
 
 	def test_hash_elements_are_mod_q
-		m = @sad.mu / 2
+		skip
+		# m = @sad.mu / 2
 
-		x = node_label(m)
-		y = node_label(m)
+		# x = column_vector(m)
+		# y = column_vector(m)
 
-		@sad.hash(x,y).each do |item|
-			assert(item < @sad.q, "#{item} is >= #{@sad.q}")
-		end
+		# @sad.hash(x,y).each do |item|
+		# 	assert(item < @sad.q, "#{item} is >= #{@sad.q}")
+		# end
 	end
 
+	def test_binary_vector
+		m = @sad.mu / 2
+
+		x = column_vector(@sad.k)
+
+		# puts
+		# puts "k: #{@sad.k}"
+		# puts "q: #{@sad.q}"
+		# puts "log q: #{Math.log2(@sad.q).ceil}"
+
+		b_size = @sad.binary_vector(x).row_size
+		expected = @sad.k * Math.log2(@sad.q).ceil
+
+		assert_equal(	expected, b_size)
+
+	end
 
 
 end

### 2012-11-06 lib/sads.rb
partial_digest method written not tested

@@ -58,7 +58,34 @@ class Sads
 	end
 
 	# private below here
-	#
+
+	def partial_digest(node_index, with_repect_to_index)
+		if node_index.length == with_repect_to_index.length
+
+			if node_index == with_repect_to_index
+
+				return "1"
+				# return Matrix.build(1, @k) { 1 }
+			else
+				raise RangeError
+			end
+		end
+
+		# is wrt_index in left or right subtree
+		#
+		# Drop the characters of wrt_index that match the node_index
+		# 	The first character of the remainder tells you
+		# 	to go left or right
+		if with_repect_to_index.sub(node_index, '')[0] == "0"
+			# Left
+			# return "L( " + partial_digest( node_index + '0' ,with_repect_to_index )
+			return @L * binary_vector(partial_digest( node_index + '0' ,with_repect_to_index ) )
+		else
+			#Right
+			# return "R( " + partial_digest( node_index + '1' ,with_repect_to_index )
+			return @R * binary_vector(partial_digest( node_index + '1' ,with_repect_to_index ) )
+		end
+	end
 
 
 

### 2012-11-06 test/test_sads.rb

@@ -92,20 +92,28 @@ class TestSads < MiniTest::Unit::TestCase
 
 	def test_binary_vector
 		m = @sad.mu / 2
-
 		x = column_vector(@sad.k)
-
-		# puts
-		# puts "k: #{@sad.k}"
-		# puts "q: #{@sad.q}"
-		# puts "log q: #{Math.log2(@sad.q).ceil}"
 
 		b_size = @sad.binary_vector(x).row_size
 		expected = @sad.k * Math.log2(@sad.q).ceil
 
 		assert_equal(	expected, b_size)
+	end
+
+	def test_partial_digest
 
 	end
 
+	def test_partial_label
+
+	end
+
+	def test_node_digest
+
+	end
+
+	def node_label
+
+	end
 
 end

### 2012-11-06 test/helper.rb
Using minitest report for colorization

@@ -11,11 +11,12 @@ rescue Bundler::BundlerError => e
 end
 
 require 'minitest/unit'
-require "minitest/reporters"
+require 'minitest/reporters'
 require 'minitest/autorun'
+require 'minitest/spec'
 MiniTest::Reporters.use! MiniTest::Reporters::DefaultReporter.new
 
-MiniTest::Reporters::SpecReporter.new
+#MiniTest::Reporters::SpecReporter.new
 
 #require 'shoulda'
 

### 2012-11-07 test/helper.rb
Using different reporter for spec and units

@@ -10,11 +10,12 @@ rescue Bundler::BundlerError => e
   exit e.status_code
 end
 
-require 'minitest/unit'
+# require 'minitest/unit'
 require 'minitest/reporters'
 require 'minitest/autorun'
-require 'minitest/spec'
-MiniTest::Reporters.use! MiniTest::Reporters::DefaultReporter.new
+# require 'minitest/spec'
+
+# MiniTest::Reporters.use! MiniTest::Reporters::DefaultReporter.new
 
 #MiniTest::Reporters::SpecReporter.new
 

### 2012-11-07 test/spec/spec_helper.rb

@@ -1,0 +1,4 @@
+require 'helper'
+require 'minitest/spec'
+
+MiniTest::Reporters.use! MiniTest::Reporters::SpecReporter.new

### 2012-11-07 test/spec/spec_sads.rb

@@ -1,5 +1,7 @@
 # sads_spec.rb
-require 'sads'
+require 'spec/spec_helper'
+require 'minitest/spec'
+require 'minitest/autorun'
 
 describe Sads do
 

### 2012-11-07 test/unit/test_sads.rb

@@ -1,4 +1,4 @@
-require 'helper'
+require 'unit/unit_helper'
 require 'sads_helper'
 require 'prime'
 

### 2012-11-07 test/unit/unit_helper.rb

@@ -1,0 +1,4 @@
+require 'helper'
+require 'minitest/unit'
+
+MiniTest::Reporters.use! MiniTest::Reporters::DefaultReporter.new

### 2012-12-17 lib/prover.rb
Added partial_labels map for storing already calculated p_labels. Added update_label for adding partial_label to an existing label when elements are added to the tree

@@ -4,7 +4,7 @@ class Prover
 	include Sads
 
 	#Data structures storing Merkle tree
-	attr_accessor :labels, :leaves, :digests
+	attr_accessor :labels, :leaves, :digests, :partial_labels
 
 	# Encryption key pair
 	attr_reader :public_key, :secret_key
@@ -59,7 +59,7 @@ class Prover
 
 		@leaves = {}
 		@labels = {}
-		@digests = {}
+		@partial_labels = {}
 
 		@root_digest = node_digest('0')
 	end # initialize
@@ -73,17 +73,22 @@ class Prover
 		# addElement has to do the following:
 		# 	* Update the leaves (or frequency value)
 		# 	* Update the labels of all affected nodes
-		# 	* Update the root digest (this may occur as part of the previous)
+		# 	* Update the root digest (this may occur as part of the previous
 
 		index = get_leaf_index(ele)
 
+		# Update leaves (frequencies)
 		if leaves.include?(index)
 			leaves[index] += 1
 		else
 			leaves[index] = 1
 		end
 
-		# update_root_digest ele
+		# Update labels for all nodes on the update path
+		get_update_path(index).each do |path_index|
+			update_label(path_index, index)
+		end
+
 	end
 
 	def removeElement(ele)
@@ -128,9 +133,17 @@ class Prover
 	def node_label(node_index)
 		 # If the digest has been computed and stored, use that, otherwise compute it
 		 # @labels[node_index] ||= calc_node_label(node_index)
-		 calc_node_label(node_index)
+		 labels[node_index] || Vector.elements( Array.new(@k * @log_q_ceil) { 0 } )
 	end
 
+
+	# Given a node index, and an index of a
+	def update_label(node_idx, wrt_index)
+		# TODO Checking wrt is a valid index with which to update node index's label
+		# puts "labels[#{node_idx}] = #{labels[node_idx]}"
+		old_label = node_label(node_idx)
+		labels[node_idx] = old_label + partial_label_wrapper(node_idx, wrt_index)
+	end
 
 	# private
 
@@ -308,11 +321,19 @@ class Prover
 
 		range_of_w.each do | leaf |
 			frequency = leaves[leaf] || 0
-			p_label = partial_label(node_index, leaf)
+			p_label = partial_label_wrapper(node_index, leaf)
 			accum += ( frequency * p_label )
 		end
 
 		return mod(accum, @q)
+	end
+
[…]

### 2012-12-17 lib/verifier.rb

@@ -39,14 +39,26 @@ class Verifier
 			parent_label  = proof[i + 1][0]
 
 			matched = check_radix_label(parent_label, hashed_digest)
-			return false if not matched
+
+			if not matched
+				# puts "check_radix_label(#{parent_label}, #{hashed_digest}) failed."
+				# puts "Checking proof[#{i}] pair against parent"
+				return false
+			end
+
 		end
 
 		# d. Check the root digest is equivalent with the known digest
 			# Check the hash of the last siblings is the root digest
 		root_child_1, root_child_2, reverse_flag = proof.last
 
-		return hash(root_child_1, root_child_2, reverse_flag) == @root_digest
+		if hash(root_child_1, root_child_2, reverse_flag) == @root_digest
+
+			return true
+		else
+			puts "Root digest did not match."
+			return false
+		end
 	end
 
 	# Given a range proof

### 2012-12-17 test/test_prover.rb

@@ -391,12 +391,11 @@ describe "Prover" do
 			all_nodes = set_of_all_node_indices(@prover)
 			all_nodes.each do |node|
 				digest = @prover.node_digest(node)
-				label = @prover.node_label(node)
+				label = @prover.calc_node_label(node)
 
-
-				# puts "Checking label: #{label}"
-				# puts "Checking digest: #{digest}"
-				# puts "q : #{@prover.q}"
+				puts "Checking label: #{label}"
+				puts "Checking digest: #{digest}"
+				puts "q : #{@prover.q}"
 
 				assert(@prover.check_radix_label(label, digest), "Label/Digest Relation Failed for node: #{node}")
 			end

### 2012-12-18 doc/final_report/draft_report.tex
Adding final report draft

--- 2012-11-04 doc/milestone_draft1.tex
@@
 \usepackage{ amssymb }
+\usepackage{graphicx}
 
@@
 
-\title{CMCS 818J Project Milestone Report}
+\title{\textbf{CMCS 818J Project Final Report}}
 \author{Behzad Koosha, Christopher Imbriano}
@@
 
-	A scheme for achieving a streaming authenticated data structure (SADS) was proposed by Papamanthou and Shi \cite{sads}.
-	The goal of this project is twofold.
-	First, we will implement an example of this data structure and second, we will build an application using the SADS primitive.
+	In this report we will be describing a scheme for achieving a streaming authenticated data structure (SADS) proposed by Papamanthou and 
+	Shi \cite{sads}. The goal of this project is twofold. First, we will be implementing an example of this data
+	structure and second, we will be building an application using the SADS primitive.
 
@@
 
-	The introduced scheme, streaming authenticated data structure, does not require any interaction between the client and the server while the stream is observed.
+	In the verifiable computation (VC) problem, a client wishes to compute a complex
+	function f on an input x. The client does not trust the server and therefore, would like to
+	be able to verify the correctness of the computation without investing too much resources.
+	The increasing use of distributed computing over the internet also makes strong servers more commonly 
+	available. The growth of outsourcable computation in the form of cloud computing has attracted a renewed
+	interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.\\
+
+	The introduced scheme, Streaming Authenticated Data Structure (SADS), does not require any interaction between the client and the server while the stream is observed.
 	Three important properties are obtained using this scheme :
@@
 	\item Expressiveness : a wide range of queries such as membership/ non-membership, successor, range search and frequencies can be supported using this scheme.
+	\end{enumerate}
+
+	The ultimate goal of the proposed scheme is to build secure protocols that enable the verification of computation
+	that is performed by an untrusted resource.\\
+
+	The rest of this paper is organized as follows. We will be describing streaming authenticated data structure 
+	goals. Some recent research in streaming verifiable computation will be discussed. We then describe the 
+	homomorphic property used in the the SADS scheme. Some definitions will be presented in the following section.
+	We present significant features of our implementation afterwards followed by presenting the performance of 
+	our investigation. Finally, we conclude our paper by providing some future plans.\\
+
+	\section{A Lattice-based Hash Tree Model}
+
+	The bottomline for the proposed scheme uses Merkle tree structure using a hash function that has the homomorphic
+	property. A way to map the outputs to the input domain considers \textit{labels} and allocates the root digest as
+	the sum of well defined functions, nameley, \textit{partial digests} applied to the leaves of the Merkle tree.
+	This will ease for the felxibility and efficiency in streaming update for the verifier and prover. \\
+
+	Each node in the tree has a label $\lambda$ and a digest \textit{d}, where a node's digest
+	could be computed uniquely from its label. Given two children labels, the digest of the parent can be 
+	calculated. Figure one represents this property. The labels are denoted by circles and digests are 
+	represented with squares. \\
+
+        \begin{figure}[h]
+        \centering
+        \includegraphics[width=100mm]{./pic/label_digest.jpg}
+        \caption[]{Labels and Digests}
+        \end{figure}
+
+	As an example, if the set of elements contains the nodes {1,4,6,7} the digest of the hash root of the tree 
+	could be expressed as : \textit{D}(1)+\textit{D}(4)+\textit{D}(6)+\textit{D}(7)
+        where \textit{D}(x) is the partial digest of the root with respect to x. \\
+
+	The hash function is calculated as follows. Given a security parameter \textit{k}, $\textit{n} = 
+	poly(\textit{k})$ and $\textit{q} $, $\textit{$\mu$}$, $\textit{$\beta$}$ outputed parameteres by the SADS 
+	algorithm, $\textbf{L}, \textbf{R} \in Z $ two (k $\times$ m) random uniform matrices in \textit{mod q}, the 
+	hashed function is defined as h(x,y)= $\textbf{L}$.x + $\textbf{R}$.y mod q. \\
+
+	\section{Properties of Digests and Labels}
+
+	\begin{enumerate}
+	\item Having a label of a node, its digest can be computed (not necessarily the other way around)
+	\item Having the labels of the two children nodes, using the hash functions, the digest of the parent
+	      could be computed.
+	\end{enumerate}
+
+	\section{Prover}
+
+	The proof of leaf node \textit{i} consists of: 
+	\begin{enumerate}
+	 \item the labels of nodes from \textit{i} to the root
+	 \item the sibling nodes to that path
+	\end{enumerate}
+
+	In figure two, the labels are presented in green (path to the root) while the sibling
+	nodes to that path are colored in blue.
+
+	\begin{figure}[h]
+        \centering
+        \includegraphics[width=100mm]{./pic/8.jpg}
+        \caption[]{Prover}
+        \end{figure}
+
+	\section{Verifier}
+
+	The verifier computes the digest for every node on the path from \textit{i} to the root in two ways, and checks
+	if they are consistent :
+
+	\begin{enumerate}
+	 \item using the \textit{labels} of its children
+	 \item from the \textit{label} of the node itself
 	\end{enumerate}
@@
 
-	We have a few frameworks in mind for this task.
-	Two of our options at the moment are EventMachine and Node.js.
-	Node.js is a "platform for building fast, scalable network applications" and has been gaining popularity in the Ruby community. http://nodejs.org/
-	EventMachine is a "fast, simple event-processing library for Ruby programs."http://rubyeventmachine.com/
+	%We have a few frameworks in mind for this task.
+	%Two of our options at the moment are EventMachine and Node.js.
+	%Node.js is a "platform for building fast, scalable network applications" and has been gaining popularity in the Ruby community. http://nodejs.org/
+	%EventMachine is a "fast, simple event-processing library for Ruby programs."http://rubyeventmachine.com/
+
+	\section{Simulation outline}
+
+	For the first stage of this project Ruby is used due to the following reasons :
+
+	\begin{enumerate}
+	\item built in support of matrix and vector operations
+	\item vibrant developer community
+	\item possibility for extension in C
+        \item numeric and visualization library
+	\end{enumerate}
 
@@
 	\subsection{Production Quality Libraries}
-	It is yet to be seen whether the existing Ruby matrix libraries, either the built in or SciRuby, will be sufficiently efficient for the needs of the SADS scheme.
-	That's not to say the current choices of technologies will not produce production quality code but, as mentioned earlier, we have been investigating some optimized libraries listed here.
+	It is yet to be seen whether the existing Ruby matrix libraries, either the built in or SciRuby, 
+	will be sufficiently efficient for the needs of the SADS scheme. That's not to say the current 
+	choices of technologies will not produce production quality code but, as mentioned earlier, we have
+	been investigating some optimized libraries listed here.
 
@@
 
-	\subsection{Semester Outlook}
-		While the ultimate goal of the project is to produce a proof-of-concept application which utilizes a SADS scheme, it is best to set manage expectations.
-		Our baseline target for this semester is to complete the first two layers described above, that is the underlying data structure and the SADS scheme primitive.
-		Having completed these two components, we can then establish time line during which the proof-of-concept application can be hashed out and developed.
-		We hope to leverage the Ruby and broader open-source community by packaging this application, or just the SADS scheme primitive, into a library for others to use and provide feedback.
+	%\subsection{Semester Outlook}
+		%While the ultimate goal of the project is to produce a proof-of-concept application which utilizes a SADS scheme, it is best to set manage expectations.
+		%Our baseline target for this semester is to complete the first two layers described above, that is the underlying data structure and the SADS scheme primitive.
+		%Having completed these two components, we can then establish time line during which the proof-of-concept application can be hashed out and developed.
+		%We hope to leverage the Ruby and broader open-source community by packaging this application, or just the SADS scheme primitive, into a library for others to use and provide feedback.
 
@@
 	\bibitem{sads} Elaine Shi, Charalampos Papamanthou \emph{Streaming Authenticated Data Structures}, Date Unknown
+	\bibitem{evsc} B. Applebaum, Y. Ishai
+and E. Kushilevitz \emph{From Secrecy to Soundness:
+Efficient Verification via Secure Computation}
 	\end{thebibliography}

### 2012-12-18 benchmark/multiply.rb
Adding preliminary banchmarking scripts

@@ -1,0 +1,22 @@
+require 'benchmark'
+require 'prover'
+
+puts "Benchmarking time to do matrix multiplications"
+Benchmark.bm(10) do |x|
+
+	(10..100).step(10) do |n|
+		(10..100).step(10) do |k|
+
+			p = Prover.new(k,n,16)
+			v = Vector.elements( Array.new(p.k * p.log_q_ceil) { rand p.q })
+
+			x.report("Matrix Mulp --  k:#{k} n:#{n}"){ 100.times { p.L * v } }
+		end
+	end
+
+
+
+
+
+
+end

### 2012-12-18 benchmark/multiply.rb
Clarified Ruby multiplication

@@ -1,22 +1,16 @@
 require 'benchmark'
 require 'prover'
 
-puts "Benchmarking time to do matrix multiplications"
-Benchmark.bm(10) do |x|
+puts "Benchmarking time to do Ruby matrix multiplications"
+Benchmark.bm(7) do |x|
 
-	(10..100).step(10) do |n|
-		(10..100).step(10) do |k|
+	(10..20).step(10) do |n|
+		(10..20).step(10) do |k|
 
 			p = Prover.new(k,n,16)
 			v = Vector.elements( Array.new(p.k * p.log_q_ceil) { rand p.q })
 
-			x.report("Matrix Mulp --  k:#{k} n:#{n}"){ 100.times { p.L * v } }
+			x.report("k:#{k} n:#{n}"){ 100.times { p.L * v } }
 		end
 	end
-
-
-
-
-
-
 end

### 2012-12-18 benchmark/generate_proof.rb
Added files as place holder for various benchmarking tasks

@@ -1,0 +1,8 @@
+require 'benchmark'
+require 'prover'
+
+puts "Benchmarking time for Prover to generate a proof"
+
+Benchmark.bm do |x|
+
+end

### 2012-12-18 benchmark/verify_proof.rb

@@ -1,0 +1,8 @@
+require 'benchmark'
+require 'verifier'
+
+puts "Benchmarking Verifier's verify proof operation"
+
+Benchmark.bm do |x|
+
+end

### 2012-12-18 misc/params.rb
Forgot to close file

@@ -25,3 +25,5 @@ f.write("k\t\t\t\tn\t\t\t\tq\t\t\t\t\tL/R dim\t\t\tEstimated Size of L or R\n")
 		f.write("#{k}\t\t#{n}\t\t#{q}\t\t#{l_rows} x #{l_cols}\t\t#{'%.0f' % l_data_size}MB\n")
 	end
 end
+
+f.close

### 2012-12-18 doc/final_report/draft_report.tex
Added tables with public key size estimations

@@
         \centering
-        \includegraphics[width=100mm]{./pic/label_digest.jpg}
+        \includegraphics[width=100mm]{./images/label_digest.jpg}
         \caption[]{Labels and Digests}
@@
         \centering
-        \includegraphics[width=100mm]{./pic/8.jpg}
+        \includegraphics[width=100mm]{./images/8.jpg}
         \caption[]{Prover}
@@
 
-	\section{Simulation outline}
+\section{ Benchmarking }
 
-	For the first stage of this project Ruby is used due to the following reasons :
+	The next stage of development is benchmarking to understand the performance of the implementation and to differentiate constraints resulting from particular design decisions from those arising from the schemes protocol itself.
+	We will focus primarily on benchmarking the timing of important operations in the scheme as well as the memory footprint of the implementations largest data structures.
 
-	\begin{enumerate}
-	\item built in support of matrix and vector operations
-	\item vibrant developer community
-	\item possibility for extension in C
-        \item numeric and visualization library
-	\end{enumerate}
+	We ran some preliminary calculations to ballpark the size of L and R the public key of the SADS scheme.
+	Recall, L and R are two matrices having elements in Z q, where q is derived from the security parameter k and the upper bound on the size of the stream, n.
+	So a rough estimate of the size of L and R would be the size of the matrices, (ie. the number of elements) multiplied by the amount of space required for each element.
+	Below is a table of these preliminary calculations.
 
-% In our work, we have generated a secret and public key depending on the size of the stream and the security parameter.
-% The next step was to generate the hash function according to each left and right leaf.
-% The hash function is computed as $ h(x,y) = Lx + Ry $ where \textbf{L, R} are selected randomly from $Z_{q}^{k \times m}$ and
-% \textbf{x,y} are vectors of small norm.
+	\begin{table}[h]
+	\centering
+		\begin{tabular}{ c | c | c | c}
+		k & n & Matrix dim & Footprint\\ \hline
+		500&10&500 x 1500&2MB\\
+		500&100&500 x 1500&2MB\\
+		500&1000&500 x 1500&2MB\\
+		500&10000&500 x 1500&2MB\\
+		500&100000&500 x 1500&3MB\\
+		500&1000000&500 x 1500&3MB\\
+		500&10000000&500 x 1500&4MB\\
+		500&100000000&500 x 1500&4MB\\
 
-%genkey procedure generates a secret and public key.  The public key includes vectors \textbf{L} and \textbf{R} as well as $q$ and $\mathcal{U}$.
+		\end{tabular}
+	\caption{ Estimation of public key size.  Constant security parameter, varying stream size upper bound }
+	\label{tab:pub-key_n}
+	\end{table}
+
+	\begin{table}[h]
+	\centering
+
+		\begin{tabular} { c | c | c | c}
+
+		k & n & Matrix dim & Footprint\\ \hline
+		1000&100000&1000 x 3000&14MB\\
+		10000&100000&10000 x 30000&1716MB\\
+		100000&100000&100000 x 300000&171661MB\\
+		1000000&100000&1000000 x 3000000&20027160MB\\
+		10000000&100000&10000000 x 30000000&2288818359MB\\	
+		\end{tabular}
+
+	\caption{Estimation of public key size.  Constant stream size upper bound , varying security parameter } 
+	\label{tab:pub-key_k}
+	\end{table}
 

### 2012-12-18 misc/params.rb
Fixed up params to format data for report (and use prover class since there's no Sads class anymore)

@@ -1,15 +1,17 @@
-require 'sads'
-s = Sads.new(0,0,0,true)
+require 'prover'
+p = Prover.new(0,0,0,true)
 f = File.open('./misc/params_bigger_steps.txt', 'w')
 
 
-f.write("k\t\t\t\tn\t\t\t\tq\t\t\t\t\tL/R dim\t\t\tEstimated Size of L or R\n")
+f.write("k\t\tn\t\tq\t\tlog_q\t\tbytes_per_entry\t\tL/R dim\t\t\tEstimated Size of L or R\n")
 
-(100..10000).step(1000) do |k|
-	(1000..10000000).step(1000000) do |n|
-		q = s.calculate_q(k, n)
+[100_000].each do |n|
+
+	(1000..10000000).step(1000) do |k|
+		q = p.calculate_q(k, n)
+		log_q = Math.log2(q).ceil
 		l_rows = k
-		l_cols = s.calculate_mu(k, q) / 2
+		l_cols = p.calculate_mu(k, q) / 2
 		l_num_elements = l_rows * l_cols
 
 		# Assuming 4 bytes per integer
@@ -17,12 +19,19 @@ f.write("k\t\t\t\tn\t\t\t\tq\t\t\t\t\tL/R dim\t\t\tEstimated Size of L or R\n")
 
 		# Each element in L can be a number up to q,
 		# 		which require the following number of bytes
-		size_per_element = Math.log2(q).ceil / 8.0
+		bytes_per_entry = (log_q / 8.0).ceil
 
 		# 1048576 bytes per megabyte
-		l_data_size = size_per_element * l_num_elements / 1048576
+		l_data_size = bytes_per_entry * l_num_elements / 1048576
+		tabs = "\t" *  ((10 - Math.log10(n)) / 2)
 
-		f.write("#{k}\t\t#{n}\t\t#{q}\t\t#{l_rows} x #{l_cols}\t\t#{'%.0f' % l_data_size}MB\n")
+		# line = [k, n, q, log_q, bytes_per_entry, "#{l_rows} x #{l_cols}", "#{'%.0f' % l_data_size}MB"].join(tabs)
+
+		line = [k, n, "#{l_rows} x #{l_cols}", "#{'%.0f' % l_data_size}MB"].join('&')
+		f.puts line + '\\\\'
+
+		# f.write("#{k}\t#{n}\t#{q}\t#{l_rows} x #{l_cols}\t#{log_q}\t#{'%.0f' % l_data_size}MB\n")
+
 	end
 end
 

### 2012-12-18 doc/final_report/draft_report.tex

@@
 
+	The estimated memory footprint of the public key matrices grows rapidly with an increasing security parameter, but slowly with increasing stream size.
+	The reason for this is that the scheme calls for matrices of size k x k log q.
+	The matrix size grows faster than k SQUARED because q also depends on k.
+	Fortunately, the current proposal is to pick a security parameter "large enough", say 500, and not to vary k over time.
+
+	Future benchmarking efforts will also measure the memory footprint of the prover over time, and the size of a proof for varying parameter sizes.
+	Additionally, we will benchmark the time required to perform the following tasks:
+
+	\begin{itemize}
+	\item Update prover state	
+	\item Proof generation
+	\item Update verifier state
+	\item Proof verification
+	\item Matrix multiplication using built-in Ruby implementation
+	\item Matrix multiplication using optimized libraries
+	\end{itemize}
+
 \section{Looking Forward}

### 2012-12-19 doc/final_report/draft_report.tex
Added to Looking Forward section.

--- 2012-12-18 doc/final_report/draft_report.tex
@@
 
-\section{ Benchmarking }
+\section{Looking Forward}
+
+	\subsection{ Benchmarking }
 
@@
 
-	We ran some preliminary calculations to ballpark the size of L and R the public key of the SADS scheme.
-	Recall, L and R are two matrices having elements in Z q, where q is derived from the security parameter k and the upper bound on the size of the stream, n.
-	So a rough estimate of the size of L and R would be the size of the matrices, (ie. the number of elements) multiplied by the amount of space required for each element.
-	Below is a table of these preliminary calculations.
+	We ran some preliminary calculations to ballpark the size of matrices L and R, the public key of the SADS scheme.
+	Recall that L and R are matrices with elements in $Z_q$, that is elements are integers of at most q and q is derived from the security parameter, k and the upper bound on the size of the stream, n.
+	So a rough estimate of the size of L and R can be calculated by multiplying the dimensions of the matrices, by the amount of space required for each element.
+	Below are two tables of estimations.
+
+	\begin{table}[h]
+	\centering
+
+		\begin{tabular} { c | c | c | c}
+
+		k & n & Matrix dim & Footprint\\ \hline
+		1000&100000&1000 x 3000&14MB\\
+		10000&100000&10000 x 30000&1716MB\\
+		100000&100000&100000 x 300000&171661MB\\
+		1000000&100000&1000000 x 3000000&20027160MB\\
+		10000000&100000&10000000 x 30000000&2288818359MB\\	
+		\end{tabular}
+
+	\caption{Estimation of public key size.  Constant stream size upper bound , varying security parameter } 
+	\label{tab:pub-key_k}
+	\end{table}
 
@@
 
-	\begin{table}[h]
-	\centering
+	The estimated memory footprint of the public key matrices grows rapidly with an increasing security parameter, but slowly with increasing stream size.
+	The reason for this is that the scheme calls for matrices of size $k  \times ( k \times \log q )$.
+	The matrix size grows faster than $k^2$ because $q$ also depends on $k$.
+	Fortunately, the current proposal is to pick one security parameter which is large enough to provide the appropriate security properties for all instances of the SADS scheme.
 
-		\begin{tabular} { c | c | c | c}
+	Future benchmarking efforts will also measure the memory footprint of the prover over time, and the size of a proof for varying parameter sizes.
+Additionally, we will benchmark the time required to perform the important updating and proof tasks for each of the prover and verifier.
 
-		k & n & Matrix dim & Footprint\\ \hline
-		1000&100000&1000 x 3000&14MB\\
-		10000&100000&10000 x 30000&1716MB\\
-		100000&100000&100000 x 300000&171661MB\\
-		1000000&100000&1000000 x 3000000&20027160MB\\
-		10000000&100000&10000000 x 30000000&2288818359MB\\	
-		\end{tabular}
+	\subsection{ Production Quality Libraries }
 
-	\caption{Estimation of public key size.  Constant stream size upper bound , varying security parameter } 
-	\label{tab:pub-key_k}
-	\end{table}
-
-\section{Looking Forward}
-
-	\subsection{Production Quality Libraries}
-	It is yet to be seen whether the existing Ruby matrix libraries, either the built in or SciRuby, 
-	will be sufficiently efficient for the needs of the SADS scheme. That's not to say the current 
-	choices of technologies will not produce production quality code but, as mentioned earlier, we have
-	been investigating some optimized libraries listed here.
+	Through our benchmarking and profiling efforts, we hope to identify and fix the bottlenecks in implementation resulting in faster and higher quality code.
+	We recognize that the built-in implementations of certain high cost operations, such as matrix multiplication may end up being the limiting factor, and as such we've identified a few options for future optimization.
+	We hesitate to invest time and effort into any of these options too soon as we know that our first-draft implementation necessarily has room for improvement, but we outline those options here.
 
 	\begin{enumerate}
+	\item \textbf{SciRuby} : A Ruby library providing efficient implementations of linear algebra operations as well as other science, engineering, and visualization packages.
+\texttt{http://sciruby.com/}
 	\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length integers, for vectors, matrices and polynomials over the integers and over the finite fields and for arbitrary precision floating point arithmetic.

### 2012-12-19 doc/final_report/draft_report.tex

@@
 
-	In this report we will be describing a scheme for achieving a streaming authenticated data structure (SADS) proposed by Papamanthou and 
-	Shi \cite{sads}. The goal of this project is twofold. First, we will be implementing an example of this data
+	While Infrastructure and Platform -as-a-Service products provide clients access to large amounts of computing power and storage capacity, the client must trust the provider with their computation.
+	It may be the case that verifying the computation requires too much resources and is infeasible for the client.
+	In this report we will discuss a streaming setting in which a client and provider both witness a stream of data.
+	In this setting the client is assumed to have limited resources in both storage and computational power, and as such, relies on the provider to serve as a database of sorts from which the client can query about the data stream.
+
+	Papamanthou and Shi \cite{sads} proposed a scheme called Streaming Authenticated Data Structures, which allows the client to verify the provider answers his or her query correctly and does not tamper with the query results.
+	The scheme has been shown to be secure and efficient requiring no communication between the client and the provider during the streaming setting.
+	The goal of this project is twofold. First, we will be implementing an example of this data
 	structure and second, we will be building an application using the SADS primitive.
@@
 
-	In the verifiable computation (VC) problem, a client wishes to compute a complex
-	function f on an input x. The client does not trust the server and therefore, would like to
-	be able to verify the correctness of the computation without investing too much resources.
-	The increasing use of distributed computing over the internet also makes strong servers more commonly 
-	available. The growth of outsourcable computation in the form of cloud computing has attracted a renewed
-	interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.\\
+	In the verifiable computation (VC) problem, a client wishes to compute a complex function f on an input x.
+	The client does not trust the server and therefore, would like to be able to verify the correctness of the computation without having to perform the computation themselves.
+	The increasing use of distributed computing over the internet also makes strong servers more commonly available.
+	The growth of outsourcable computation in the form of cloud computing has attracted a renewed interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.
 
@@
 	\end{enumerate} 
+	The ultimate goal of the proposed scheme is to build secure protocols that enable the verification of computation 	that is performed by an untrusted resource.
 
-	The ultimate goal of the proposed scheme is to build secure protocols that enable the verification of computation
-	that is performed by an untrusted resource.\\
-
-	The rest of this paper is organized as follows. We will be describing streaming authenticated data structure 
-	goals. Some recent research in streaming verifiable computation will be discussed. We then describe the 
-	homomorphic property used in the the SADS scheme. Some definitions will be presented in the following section.
-	We present significant features of our implementation afterwards followed by presenting the performance of 
-	our investigation. Finally, we conclude our paper by providing some future plans.\\
+	The rest of this paper is organized as follows. We will be describing streaming authenticated data structure goals.
+	Some recent research in streaming verifiable computation will be discussed.
+	We then describe the homomorphic property used in the the SADS scheme. Some definitions will be presented in the following section.
+	We present significant features of our implementation afterwards followed by presenting the performance of our investigation.
+	Finally, we conclude our paper by providing some future plans.
 
@@
 
-	The bottomline for the proposed scheme uses Merkle tree structure using a hash function that has the homomorphic
-	property. A way to map the outputs to the input domain considers \textit{labels} and allocates the root digest as
-	the sum of well defined functions, nameley, \textit{partial digests} applied to the leaves of the Merkle tree.
-	This will ease for the felxibility and efficiency in streaming update for the verifier and prover. \\
+	The bottomline for the proposed scheme uses Merkle tree structure using a hash function that has the homomorphic property.
+	A way to map the outputs to the input domain considers \textit{labels} and allocates the root digest as	the sum of well defined functions, nameley, \textit{partial digests} applied to the leaves of the Merkle tree.
+	This will ease for the felxibility and efficiency in streaming update for the verifier and prover.
 
-	Each node in the tree has a label $\lambda$ and a digest \textit{d}, where a node's digest
-	could be computed uniquely from its label. Given two children labels, the digest of the parent can be 
-	calculated. Figure one represents this property. The labels are denoted by circles and digests are 
-	represented with squares. \\
+	Each node in the tree has a label $\lambda$ and a digest \textit{d}, where a node's digest could be computed uniquely from its label.
+	Given two children labels, the digest of the parent can be calculated. 
+	Figure one represents this property. The labels are denoted by circles and digests are 
+	represented with squares.
 
@@
 	could be expressed as : \textit{D}(1)+\textit{D}(4)+\textit{D}(6)+\textit{D}(7)
-        where \textit{D}(x) is the partial digest of the root with respect to x. \\
+        where \textit{D}(x) is the partial digest of the root with respect to x.
 
 	The hash function is calculated as follows. Given a security parameter \textit{k}, $\textit{n} = 
-	poly(\textit{k})$ and $\textit{q} $, $\textit{$\mu$}$, $\textit{$\beta$}$ outputed parameteres by the SADS 
+	poly(\textit{k})$ and $\textit{q} $, $\textit{$\mu$}$, $\textit{$\beta$}$ outputed parameters by the SADS 
 	algorithm, $\textbf{L}, \textbf{R} \in Z $ two (k $\times$ m) random uniform matrices in \textit{mod q}, the 
-	hashed function is defined as h(x,y)= $\textbf{L}$.x + $\textbf{R}$.y mod q. \\
+	hashed function is defined as h(x,y)= $\textbf{L}$.x + $\textbf{R}$.y mod q.
 
@@
 		\item Having a label of a node, its digest can be computed (not necessarily the other way around)
-	\item Having the labels of the two children nodes, using the hash functions, the digest of the parent
-	      could be computed.
+		\item Having the labels of the two children nodes, using the hash functions, the digest of the parent could be computed.
 	\end{enumerate}
@@
 
-	In figure two, the labels are presented in green (path to the root) while the sibling
-	nodes to that path are colored in blue.
+	In figure two, the labels are presented in green (path to the root) while the sibling nodes to that path are colored in blue.
 
@@
 
-	The verifier computes the digest for every node on the path from \textit{i} to the root in two ways, and checks
-	if they are consistent :
+	The verifier computes the digest for every node on the path from \textit{i} to the root in two ways, and checks if they are consistent :
 
@@
 
-	The description of the SADS scheme reads much like a protocol, focusing on the behaviour of each of the prover and verifier during a streaming phase and a query phase.
+	The description of the SADS scheme reads much like a protocol, focusing on the behavior of each of the prover and verifier during a streaming phase and a query phase.
 	Before tackling the problem at the protocol level however, we must first implement the basic data structures each of the verifier and prover will need to perform their roles.

### 2012-12-19 doc/final_report/draft_report.tex

@@
 	In the verifiable computation (VC) problem, a client wishes to compute a complex function f on an input x.
-	The client does not trust the server and therefore, would like to be able to verify the correctness of the computation without having to perform the computation themselves.
-	The increasing use of distributed computing over the internet also makes strong servers more commonly available.
-	The growth of outsourcable computation in the form of cloud computing has attracted a renewed interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.
+	The client does not trust the server and therefore, would like to be able to verify the correctness of the computation without having to perform the entire computation themselves.
+	The increasing use of distributed computing over the internet also makes commoditizes powerful servers.
+	The growth of outsourcing computation in the form of cloud computing has attracted a renewed interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.
 
@@
 	\begin{enumerate}[a.]
-		\item Independence of prover and verifier : the prover and the verifier update their status independently and with no interaction.
+		\item Independence of prover (server) and verifier(client) : the prover and the verifier update their internal state without interaction.
 		\item Efficiency : It can be shown that an exponential improvement for many commonly queries in the prover's running time can be achieved.
@@
 	\end{enumerate} 
-	The ultimate goal of the proposed scheme is to build secure protocols that enable the verification of computation 	that is performed by an untrusted resource.
+	The ultimate goal of the proposed scheme is to implement the SADS scheme that enable the verification of computation performed by an untrusted resource.
 
@@
 
-	\section{A Lattice-based Hash Tree Model}
+	\section{ Achieving independence through \\ a Lattice-based Hash Tree }
 
-	The bottomline for the proposed scheme uses Merkle tree structure using a hash function that has the homomorphic property.
-	A way to map the outputs to the input domain considers \textit{labels} and allocates the root digest as	the sum of well defined functions, nameley, \textit{partial digests} applied to the leaves of the Merkle tree.
-	This will ease for the felxibility and efficiency in streaming update for the verifier and prover.
+	The core concept that enables the SADS scheme to satisfy the desired goals stated above is a lattice-based hash tree, also known as a Merkle tree.
+	In traditional Merkle trees, a hash function such as SHA-1 might be used to determine the parent of any two children.
+	SHA-1 and similar hashes do not have homomorphic properties which would enable computation over two hashes.
+	In the case of the SADS scheme, the hash used is one with homomorphic properties allowing each node in the tree to be represented as a sum of parts, including the root of the tree.
 
-	Each node in the tree has a label $\lambda$ and a digest \textit{d}, where a node's digest could be computed uniquely from its label.
-	Given two children labels, the digest of the parent can be calculated. 
-	Figure one represents this property. The labels are denoted by circles and digests are 
-	represented with squares.
+	This homomorphic hash function will map what are called \textit{labels} to \textit{digests}, where labels and digests are different representations of the same value for each node.
+	As in a traditional Merkle tree, each node with children derives its value from the hash of its children.
+	In this case, a parent's digest is exactly the hash of its children's labels.
+	This concept is depicted in figure \ref{fig:labels-and-digests}.	
+	Each node in the tree has a label $\lambda$ and a digest \textit{d}.
 
@@
         \caption[]{Labels and Digests}
+        \label{fig:labels-and-digests}
         \end{figure}
 
-	As an example, if the set of elements contains the nodes {1,4,6,7} the digest of the hash root of the tree 
-	could be expressed as : \textit{D}(1)+\textit{D}(4)+\textit{D}(6)+\textit{D}(7)
-        where \textit{D}(x) is the partial digest of the root with respect to x.
+	The key homomorphic property, which will only be briefly described here, allows every node in the tree to be represented by a sum of representations of all its children.
+	Each of those children can be represented as a sum of their children and so on.
+	In this way, even the root of the tree can be represented by a sum of representations of the leaves of the tree.
+	The representation being alluded to is called a \textit{partial digest} and is calculated for a node, \textit{with respect to} some node in that parent's subtree.
 
-	The hash function is calculated as follows. Given a security parameter \textit{k}, $\textit{n} = 
-	poly(\textit{k})$ and $\textit{q} $, $\textit{$\mu$}$, $\textit{$\beta$}$ outputed parameters by the SADS 
-	algorithm, $\textbf{L}, \textbf{R} \in Z $ two (k $\times$ m) random uniform matrices in \textit{mod q}, the 
-	hashed function is defined as h(x,y)= $\textbf{L}$.x + $\textbf{R}$.y mod q.
+	For example, suppose a tree is to represent a set of elements drawn from a universe on integers in $0 \leq x \leq 7$.
+	If then we want to represent the root of such a tree in which the set of nodes \{1, 4, 6, 7\} are present, then digest of the root of the tree can be expressed as a sum of partial digests of the root, with respect to the present elements.
+	That is $\textit{D}(root, 1)+\textit{D}(root, 4)+\textit{D}(root, 6)+\textit{D}(root, 7)$ where $\textit{D}(root, x)$ is the partial digest of the root with respect to x.
+	This type of hash function is what allows the client (verifier) to maintain a correct running digest of all the stream elements seen, regardless of how large the tree is or how many stream elements have been witnessed.
+
+	For the readers knowledge, we describe the hash function, but will not go into detail.  The proof of its security and correctness properties can be found in \cite{sads}.
+	The hash function is:
+
+	\begin{equation}
+		h(x,y)= \textbf{L} x + \textbf{R}y
+	\end{equation}
+	where $x, y$ are labels of nodes, $\textbf{L}, \textbf{R}$ are matrices performing the role of a public key in the scheme and $h(x,y)$ is a digest.
 

### 2012-12-20 doc/final_report/draft_report.tex

@@
 
-	\section{Prover}
+\section{ Participants }
 
-	The proof of leaf node \textit{i} consists of: 
-	\begin{enumerate}
-		\item the labels of nodes from \textit{i} to the root
-		\item the sibling nodes to that path
-	\end{enumerate}
+	\subsection{ Verifier and Prover }
 
-	In figure two, the labels are presented in green (path to the root) while the sibling nodes to that path are colored in blue.
+		As mentioned previously, the verifier (or client) is a low resource computational device.
+		It is expected that the verifier will not have sufficient storage capacity to store every element of a stream witnessed and as such it has no means to correctly answer queries based on the stream elements.
+		To accomplish these tasks, the verifier will rely on a more powerful device.
+		The verifier will however have the means to storage a digest, a small representation of all the elements it has seen, with the goal being that it will be able to update its local state efficiently enough so as to not disrupt viewing elements of the stream.
+		So, the verifier's only operation during the streaming phase will be to update its local state with some representation of each element of the stream.
+		This representation will be the partial digest of the root with respect to the witnessed element as previously discussed.
+
+		The prover (server) on the other hand must act somewhat as a database.
+		Internally, the prover maintains a representation of the stream elements witnessed such that it can answer queries from the verifier.
+		That internal representation is the hash tree mentioned earlier which allows for a node to be update without rehashing its children.
+		The verifier exploits this property to maintain and update the root digest of the tree without communication with the prover.
+
+		\subsection{ Query Proof }
+
+		During or after the streaming phase, the verifier may wish to query the prover (server) for the membership, successor, range, frequency etc of a universe element that maybe or may not have been witnessed.
+		The prover is expected to provide and answer to the query as well as a proof that can be efficiently verified by the verifier. 
+		Figure \ref{fig:proof} in a visual representation of a membership proof for the green node.
+		The set of green nodes comprises the \textit{update path} for the green leaf node
+		The set of blue nodes are the \textit{siblings} of the nodes on the update path.
+		The prover then responds to a membership query with the computed answer as well as the set of sibling pairs which the verifier can then verify.
+
+		Verification is done in two parts.
+		First, for each sibling pair is hashed to obtain a digest of their parent.
+		That calculated digest must be an equivalent representation of the label in the next pair representing this pairs parent.
+		Lastly, the digest produced by hashing the last pair of children must be equivalent to the running digest the verifier has been maintaining.
+		Shi and Papamanthou \cite{dads} describe the correctness properties in much more detail.
 
@@
 		\caption[]{Prover}
+		\label{fig:proof}
 		\end{figure}
 
-	\section{Verifier}
+\section{ Implementation }
 
-	The verifier computes the digest for every node on the path from \textit{i} to the root in two ways, and checks if they are consistent :
+	Implementing the SADS scheme was organized as follows.
+	First to be implemented was the data underlying structure used by the prover and mechanisms used by both prover and verifier to support the SADS scheme.
+	Then the data structures and mechanism to support the query phase was added.
+	Finally, these two parts were packaged such that another project may leverage the SADS scheme as a library primitive and build applications.
 
-	\begin{enumerate}
-	 \item using the \textit{labels} of its children
-	 \item from the \textit{label} of the node itself
-	\end{enumerate}
+	Our language of choice for this first stage of the project was Ruby.
+	Ruby's main advantage was the ease of rapid prototyping as well as the vast collection of libraries to enhance development and functionality available.
+	Specifically, the main component of the SADS scheme is the hash function which require matrix multiplication support.
+	Ruby's standard library includes Vector and Matrix arithmetic operations which were sufficient for prototype scale implementations.
+	Ruby is also a forward looking choice as we still have the option to create bindings for C and C++ code, should we require more efficient linear algebra libraries.
 
-	\section{Our Approach}
-
-	The description of the SADS scheme reads much like a protocol, focusing on the behavior of each of the prover and verifier during a streaming phase and a query phase.
-	Before tackling the problem at the protocol level however, we must first implement the basic data structures each of the verifier and prover will need to perform their roles.
-	This will be our starting point.
-
-	The road map for this project is as follows:
-	\begin{enumerate}
-	\item Implement underlying Merkle tree with homomorphic hash function
-	\item Implement the SADS scheme
-	\item Develop proof-of-concept application using the SADS scheme implementation
-	\end{enumerate}
+	The implementation in its current state as of this writing consists of two Ruby classes and one Ruby module.
+	The classes are, not surprising, Prover and Verifer and the module is named Sads.
+	Ruby has a feature called \textit{mixing} in which a module can be "mixed into" a class.
+	Mixing in a module has the effect of include all the modules fields and methods into the class definition.
 
@@
 		Instead, we've decided to use an interpreted language and the simple matrix operations that come with the standard libraries to start.
-
-	Our language of choice for this first stage of the project is Ruby. Ruby has a number of advantages including:
-
-	\begin{enumerate}
-	\item Built-in support of matrix and vector operations
-	\item Vibrant developer community that supports both new and experienced developers through learning resources, tool development, etc...
-	\item Ability to write compiled extensions in C
-	\item Numeric and visualization library, SciRuby
-	\end{enumerate}
 

### 2012-12-20 doc/final_report/draft_report.tex
Rough draft finished

@@
 	The implementation in its current state as of this writing consists of two Ruby classes and one Ruby module.
-	The classes are, not surprising, Prover and Verifer and the module is named Sads.
-	Ruby has a feature called \textit{mixing} in which a module can be "mixed into" a class.
+	The classes are, not surprising, Prover and Verifier and the module is named Sads.
+	Ruby has a feature called \textit{mixins} in which a module can be "mixed into" a class.
 	Mixing in a module has the effect of include all the modules fields and methods into the class definition.
+	The Sads module contains any pieces of functionality required by both the Prover and the Verifier - the definition of the hash function for example.
+	However since an instantiation of a Sads would not make sense, module mix in is used instead of class inheritance.
+	The Prover and Verifier classes then mix in the functionality defined in Sads augmenting the Prover-only or Verifier-only functionality defined in their respective classes.
+		Future iterations of this library will include a stronger focus on object-oriented design especially surrounding the nodes of the tree stored by the prover, the proofs generated per query requests, and abstractions surrounding the matrix multiplication so various linear algebra implementations can be tested.
 
-	\subsection{Underlying Data Structures}
+	The next stage was to implement streaming setting in which the Prover and Verifier operate independently while witnessing a common stream of data.
+	We wrote a very simple script to simulate such a scenario.
+	This script was created as a new project, also using Ruby, and installed the SADS library via the RubyGems command 'gem install --local'.
+	We wanted to ensure that the steps we took to use the SADS library in another project were the same or close to what other developers might do when using our library.
+	The script takes three parameters to initialize the entire protocol, the security parameter, the upper bound on the size of the stream, and the size of the universe of stream elements.
+	The Prover can be initialized from just these three parameter.
+	The Verifier is then initialized with parameters provided to it from the Prover, similar to a client-server initialization wherein a client requests a session from a web server and can only proceed once the server grants the client the credentials.
+	Finally in this small example, the user is then able to run commands such as "add" which simulates the Prover and Verifier both witnessing some stream element or "query" which simulates a membership query from the Verifier to the Prover.
 
-		A crucial component of the SADS algorithm relies on matrix multiplication, a relatively expensive computational operation.
-		When formulating our plan of action for this project, we were recommended a few linear algebra libraries, which will be discussed in a later section.
-		While we recognize the importance of fast matrix operations and parallelism for production quality code, we thought that the added complexity, as well as unknowns with respect to implementation details, would be prohibitive.
-		Instead, we've decided to use an interpreted language and the simple matrix operations that come with the standard libraries to start.
-
-	\subsection{SADS Scheme}
-
-	Once the underlying data structure is complete, the SADS scheme will use it as a primitive.
-	Though one of the scheme's strengths is the independence of the prover and verifier during the streaming setting, the two parties will necessarily need to communicate during the query phase.
-	Additionally, there may be communication between the two parties to initiate the streaming setting.
-	For these reasons, we anticipate creating an application which will require network communication as it will likely be the case that the verifier and prover are not collocated.
-
-	%We have a few frameworks in mind for this task.
-	%Two of our options at the moment are EventMachine and Node.js.
-	%Node.js is a "platform for building fast, scalable network applications" and has been gaining popularity in the Ruby community. http://nodejs.org/
-	%EventMachine is a "fast, simple event-processing library for Ruby programs."http://rubyeventmachine.com/
+	The final stage is to implement the simulation across distributed computing infrastructure, that is a Prover and Verifier non-co-located.
+	We imagine at least two possibilities for this implementation.
+	One possibility could be a web service could be created that plays matchmaker between "needy" clients requesting prover services and "generous" clients wishing to act as a prover.
+	Another possibility, similar to the first, is one in which the web service itself acts as a Prover for any client requesting service.
+	In either of these scenarios we imagine deploying a server application possibly build in EventMachine or Node.js or some similar framework.
+	Some proposed applications include monitoring and verifying of bandwidth charges from an ISP or continuous sensor data.
 
 \section{Looking Forward}
+
+	\subsection{ Production Quality Libraries }
+
+	Through our benchmarking and profiling efforts, we hope to identify and fix the bottlenecks in implementation resulting in faster and higher quality code.
+	We recognize that the built-in implementations of certain high cost operations, such as matrix multiplication may end up being the limiting factor, and as such we've identified a few options for future optimization.
+	We hesitate to invest time and effort into any of these options too soon as we know that our first-draft implementation necessarily has room for improvement, but we outline those options here.
+
+	\begin{enumerate}
+	\item \textbf{SciRuby} : A Ruby library providing efficient implementations of linear algebra operations as well as other science, engineering, and visualization packages.
+\texttt{http://sciruby.com/}
+	\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length integers, for vectors, matrices and polynomials over the integers and over the finite fields and for arbitrary precision floating point arithmetic.
+\texttt{http://shoup.net/ntl/}
+	\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of the important concrete classes of structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, module theory and the theory of algebras.
+	In addition, certain families of structures from algebraic geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
+	\end{enumerate}
 
@@
 
-	We ran some preliminary calculations to ballpark the size of matrices L and R, the public key of the SADS scheme.
-	Recall that L and R are matrices with elements in $Z_q$, that is elements are integers of at most q and q is derived from the security parameter, k and the upper bound on the size of the stream, n.
-	So a rough estimate of the size of L and R can be calculated by multiplying the dimensions of the matrices, by the amount of space required for each element.
+	We ran some preliminary calculations to ballpark the size of matrices \textbf{L} and \textbf{R}, the public key of the SADS scheme.
+	Recall that \textbf{L} and \textbf{R} are matrices with elements in $Z_q$, that is elements are integers of at most $q$, which is itself derived from the security parameter, $k$ and the upper bound on the size of the stream, $n$.
+	So a rough estimate of the size of \textbf{L} and \textbf{R} can be calculated by multiplying the dimensions of the matrices, by the amount of space required for each element.
 	Below are two tables of estimations.
@@
 
-	\caption{Estimation of public key size.  Constant stream size upper bound , varying security parameter } 
-	\label{tab:pub-key_k}
+	\caption{Estimation of public key size.  Constant stream size upper bound, varying security parameter } 
+	\label{tab:pub-key-k}
 	\end{table}
@@
 	\caption{ Estimation of public key size. Constant security parameter, varying stream size upper bound }
-	\label{tab:pub-key_n}
+	\label{tab:pub-key-n}
 	\end{table}
@@
 Additionally, we will benchmark the time required to perform the important updating and proof tasks for each of the prover and verifier.
-
-	\subsection{ Production Quality Libraries }
-
-	Through our benchmarking and profiling efforts, we hope to identify and fix the bottlenecks in implementation resulting in faster and higher quality code.
-	We recognize that the built-in implementations of certain high cost operations, such as matrix multiplication may end up being the limiting factor, and as such we've identified a few options for future optimization.
-	We hesitate to invest time and effort into any of these options too soon as we know that our first-draft implementation necessarily has room for improvement, but we outline those options here.
-
-	\begin{enumerate}
-	\item \textbf{SciRuby} : A Ruby library providing efficient implementations of linear algebra operations as well as other science, engineering, and visualization packages.
-\texttt{http://sciruby.com/}
-	\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length integers, for vectors, matrices and polynomials over the integers and over the finite fields and for arbitrary precision floating point arithmetic.
-\texttt{http://shoup.net/ntl/}
-	\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of the important concrete classes of structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, module theory and the theory of algebras.
-	In addition, certain families of structures from algebraic geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
-	\end{enumerate}
 
@@
 	\bibitem{sads} Elaine Shi, Charalampos Papamanthou \emph{Streaming Authenticated Data Structures}, Date Unknown
-	\bibitem{evsc} B. Applebaum, Y. Ishai
-and E. Kushilevitz \emph{From Secrecy to Soundness:
-Efficient Verification via Secure Computation}
+	\bibitem{evsc} B. Applebaum, Y. Ishai and E. Kushilevitz \emph{From Secrecy to Soundness: Efficient Verification via Secure Computation}
 	\end{thebibliography}

### 2012-12-20 doc/final_report/draft_report.tex
Edits to Abstract

@@
 
-\title{\textbf{CMCS 818J Project Final Report}}
+\title{\textbf{CMCS 818J Project Final Report} \\ Implementing Streaming Authenticated Data Structures}
 \author{Behzad Koosha, Christopher Imbriano}
@@
 	While Infrastructure and Platform -as-a-Service products provide clients access to large amounts of computing power and storage capacity, the client must trust the provider with their computation.
-	It may be the case that verifying the computation requires too much resources and is infeasible for the client.
-	In this report we will discuss a streaming setting in which a client and provider both witness a stream of data.
-	In this setting the client is assumed to have limited resources in both storage and computational power, and as such, relies on the provider to serve as a database of sorts from which the client can query about the data stream.
+	It may be the case that verifying, by replication of the computation, requires too many resources and is infeasible for the client.
+	In this report we will discuss a scheme for computation verification in a streaming setting; that is, one in which a client and provider both witness a stream of data.
+	In this setting the client is assumed to have limited resources in both storage and computational power, and must rely on the provider to serve as a database of sorts from which the client can query about the data stream.
 
-	Papamanthou and Shi \cite{sads} proposed a scheme called Streaming Authenticated Data Structures, which allows the client to verify the provider answers his or her query correctly and does not tamper with the query results.
+	Shi and Papamanthou \cite{sads} proposed a scheme called Streaming Authenticated Data Structures which allows a client to verify the correctness and completeness of answers to queries given by an untrusted party.
 	The scheme has been shown to be secure and efficient requiring no communication between the client and the provider during the streaming setting.
-	The goal of this project is twofold. First, we will be implementing an example of this data
-	structure and second, we will be building an application using the SADS primitive.
+	The goal of this project is twofold: first, produce this data structure as a library primitive and second produce an application demonstrating the library.
+	So far we have implemented this type of data structure as a Ruby library and written a script to demonstrate the basic properties second.
+	We also outline improvements and tasks required to realize an application using the SADS primitive.
 
@@
 
-	In the verifiable computation (VC) problem, a client wishes to compute a complex function f on an input x.
-	The client does not trust the server and therefore, would like to be able to verify the correctness of the computation without having to perform the entire computation themselves.
-	The increasing use of distributed computing over the internet also makes commoditizes powerful servers.
-	The growth of outsourcing computation in the form of cloud computing has attracted a renewed interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.
+	A verifiable computation (VC) is one in which computation can be performed on behalf of a requester, or client, by an untrusted resource but can be verified to be correct by the client.
+	In this scenario, the client usually cannot, or wishes not to, perform the entire computation themselves and thus needs a way to verify the computation efficiently. 
+	The growth of outsourced computation in the form of cloud computing has attracted a strong interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.
 
-	The introduced scheme, Streaming Authenticated Data Structure (SADS), does not require any interaction between the client and the server while the stream is observed.
+	The Streaming Authenticated Data Structure scheme (SADS) does not require any interaction between the client and the server while the stream is observed.
 	Three important properties are obtained using this scheme :

### 2012-12-20 doc/final_report/draft_report.tex
Edits Section 2

@@
 
-	A verifiable computation (VC) is one in which computation can be performed on behalf of a requester, or client, by an untrusted resource but can be verified to be correct by the client.
-	In this scenario, the client usually cannot, or wishes not to, perform the entire computation themselves and thus needs a way to verify the computation efficiently. 
-	The growth of outsourced computation in the form of cloud computing has attracted a strong interest in the VC problem, and a considerable amount of research has been devoted to these problems \cite{evsc}.
+	A verifiable computation (VC) is one in which computation can be performed by a server on behalf of a client where the server is an untrusted resource.
+	In this scenario, the client usually cannot, or wishes not to, perform the entire computation themselves but requires a means of verifying the computation efficiently.
+	The growth of outsourced computation in the form of cloud computing has attracted a strong interest in the VC problem.
+	In the streaming setting, many solutions to this problem have been proposed but so far have required inefficient protocols requiring the client and server to communicate updates as each stream element arrives.
+	This communication requirement is prohibitively expensive in certain streaming applications where elements arrive in rapid succession, or are continuous.
 
-	The Streaming Authenticated Data Structure scheme (SADS) does not require any interaction between the client and the server while the stream is observed.
+	The Streaming Authenticated Data Structure scheme (SADS) \cite{dads} does not require any interaction between the client and the server while the stream is observed.
 	Three important properties are obtained using this scheme :
 	\begin{enumerate}[a.]
-		\item Independence of prover (server) and verifier(client) : the prover and the verifier update their internal state without interaction.
-		\item Efficiency : It can be shown that an exponential improvement for many commonly queries in the prover's running time can be achieved.
-		\item Expressiveness : a wide range of queries such as membership/ non-membership, successor, range search and frequencies can be supported using this scheme.
+		\item Independence of prover (server) and verifier (client) : the prover and the verifier can update internal state without interaction.
+		\item Efficiency : SADS demonstrates an exponential improvement in server's running time for many common query types.
+		\item Expressiveness : a wide range of queries such as (non-)membership, successor, range search and frequencies can be supported
 	\end{enumerate} 
@@
 
-	The rest of this paper is organized as follows. We will be describing streaming authenticated data structure goals.
-	Some recent research in streaming verifiable computation will be discussed.
-	We then describe the homomorphic property used in the the SADS scheme. Some definitions will be presented in the following section.
-	We present significant features of our implementation afterwards followed by presenting the performance of our investigation.
-	Finally, we conclude our paper by providing some future plans.
+	The rest of this paper is organized as follows. 
+	Section \ref{sec:tree} will discuss the key data structure use to achieve the SADS goals.
+	Section \ref{sec:parts} will discuss the key components in the SADS scheme.
+	Section \ref{sec:impl} will discuss details of our implementation \ref{sec:forward} will discuss preliminary measurements and future improvements.
 
-\section{ Achieving independence through \\ a Lattice-based Hash Tree }
+\section{ Achieving Independence through \\ a Lattice-based Hash Tree }
+\label{sec:tree}
 
@@
 \section{ Participants }
+\label{sec:parts}
 
@@
 \section{ Implementation }
+\label{sec:impl}
 
@@
 \section{Looking Forward}
-
+\label{sec:forward}
 	\subsection{ Production Quality Libraries }
@@
 	\bibitem{sads} Elaine Shi, Charalampos Papamanthou \emph{Streaming Authenticated Data Structures}, Date Unknown
-	\bibitem{evsc} B. Applebaum, Y. Ishai and E. Kushilevitz \emph{From Secrecy to Soundness: Efficient Verification via Secure Computation}
+%	\bibitem{evsc} B. Applebaum, Y. Ishai and E. Kushilevitz \emph{From Secrecy to Soundness: Efficient Verification via Secure Computation}
 	\end{thebibliography}

### 2012-12-20 doc/final_report/draft_report.tex

@@
 
-	The Streaming Authenticated Data Structure scheme (SADS) \cite{dads} does not require any interaction between the client and the server while the stream is observed.
+	The Streaming Authenticated Data Structure scheme (SADS) \cite{sads} does not require any interaction between the client and the server while the stream is observed.
 	Three important properties are obtained using this scheme :

### 2012-12-20 doc/final_report/draft_report.tex
Up to section 5 editied

@@
 	Section \ref{sec:parts} will discuss the key components in the SADS scheme.
-	Section \ref{sec:impl} will discuss details of our implementation \ref{sec:forward} will discuss preliminary measurements and future improvements.
+	Section \ref{sec:impl} will discuss details of our implementation and section \ref{sec:forward} will discuss preliminary measurements and outline future improvements.
 
@@
 
-	The core concept that enables the SADS scheme to satisfy the desired goals stated above is a lattice-based hash tree, also known as a Merkle tree.
-	In traditional Merkle trees, a hash function such as SHA-1 might be used to determine the parent of any two children.
-	SHA-1 and similar hashes do not have homomorphic properties which would enable computation over two hashes.
+	The core concept that enables the SADS scheme to satisfy its desired goals is a lattice-based hash tree, also known as a Merkle tree.
+	In traditional Merkle trees a hash function, such as SHA-1, is used to compute the value of any internal node of the tree.
+	SHA-1, and similar hashes, are not homomorphic hash functions; that is, the hash of the sum of two elements is not necessarily the sum of the hash of the individual elements.
+	For example, suppose there is a hash $h(x, y)$, a homomorphic hash would satisfy the following relation:
+	\begin{equation}
+		h(x + y) = h(x) + h(y)
+	\end{equation}
+	The key to the SADS scheme is Merkle tree using a homomorphic hash function which allows 
 	In the case of the SADS scheme, the hash used is one with homomorphic properties allowing each node in the tree to be represented as a sum of parts, including the root of the tree.
 
-	This homomorphic hash function will map what are called \textit{labels} to \textit{digests}, where labels and digests are different representations of the same value for each node.
-	As in a traditional Merkle tree, each node with children derives its value from the hash of its children.
+	The homomorphic hash function in the SADS scheme maps what are called \textit{labels} to \textit{digests}, where labels and digests are different representations of the same value for each node.
+	As in a traditional Merkle tree, each internal node derives its value from the hash of its children.
 	In this case, a parent's digest is exactly the hash of its children's labels.
@@
 
-	The key homomorphic property, which will only be briefly described here, allows every node in the tree to be represented by a sum of representations of all its children.
+	The crucial property of the hash function in the SADS scheme, which will only be briefly described here, allows every node in the tree to be represented by a sum of representations of its children.
 	Each of those children can be represented as a sum of their children and so on.
 	In this way, even the root of the tree can be represented by a sum of representations of the leaves of the tree.
-	The representation being alluded to is called a \textit{partial digest} and is calculated for a node, \textit{with respect to} some node in that parent's subtree.
+	The representation being alluded to is called a \textit{partial digest} and is calculated for a particular node \textit{with respect to} some descendant in that nodes subtree.
 
-	For example, suppose a tree is to represent a set of elements drawn from a universe on integers in $0 \leq x \leq 7$.
-	If then we want to represent the root of such a tree in which the set of nodes \{1, 4, 6, 7\} are present, then digest of the root of the tree can be expressed as a sum of partial digests of the root, with respect to the present elements.
-	That is $\textit{D}(root, 1)+\textit{D}(root, 4)+\textit{D}(root, 6)+\textit{D}(root, 7)$ where $\textit{D}(root, x)$ is the partial digest of the root with respect to x.
+	For example, suppose a tree is to represent a set of elements drawn from a universe of integers in the range $0 \leq x \leq 7$.
+	To represent the root of such a tree in which the set of nodes \{1, 4, 6, 7\} are present, then digest of the root of the tree can be expressed as a sum of partial digests of the root, with respect to the present elements.
+	That is,
+	\begin{equation}
+		\textit{D}(root, 1)+\textit{D}(root, 4)+\textit{D}(root, 6)+\textit{D}(root, 7)
+	\end{equation} 
+	where $\textit{D}(root, x)$ is the partial digest of the root with respect to x.
 	This type of hash function is what allows the client (verifier) to maintain a correct running digest of all the stream elements seen, regardless of how large the tree is or how many stream elements have been witnessed.
@@
 	For the readers knowledge, we describe the hash function, but will not go into detail.  The proof of its security and correctness properties can be found in \cite{sads}.
-	The hash function is:
+	The hash function is given by,
 
@@
 	\end{equation}
-	where $x, y$ are labels of nodes, $\textbf{L}, \textbf{R}$ are matrices performing the role of a public key in the scheme and $h(x,y)$ is a digest.
-
-\section{Properties of Digests and Labels}
-
-	\begin{enumerate}
-		\item Having a label of a node, its digest can be computed (not necessarily the other way around)
-		\item Having the labels of the two children nodes, using the hash functions, the digest of the parent could be computed.
-	\end{enumerate}
+	where $x, y$ are labels of nodes, represented as vectors, $\textbf{L}, \textbf{R}$ are matrices performing the role of a public key in the scheme and $h(x,y)$ is a digest also represented as a vector.
 
@@
 		As mentioned previously, the verifier (or client) is a low resource computational device.
-		It is expected that the verifier will not have sufficient storage capacity to store every element of a stream witnessed and as such it has no means to correctly answer queries based on the stream elements.
+		It is expected that the verifier will not have sufficient storage capacity to store every element of a stream witnessed and therefore cannot correctly answer queries over the data stream elements.
 		To accomplish these tasks, the verifier will rely on a more powerful device.
-		The verifier will however have the means to storage a digest, a small representation of all the elements it has seen, with the goal being that it will be able to update its local state efficiently enough so as to not disrupt viewing elements of the stream.
-		So, the verifier's only operation during the streaming phase will be to update its local state with some representation of each element of the stream.
-		This representation will be the partial digest of the root with respect to the witnessed element as previously discussed.
+		The verifier will however have the means to store a digest, a small representation of all the elements it has seen.
+		It will also be able to efficiently update its local state so as to not disrupt viewing elements of the stream.
+		The verifier's only operation during the streaming phase will be to update its local state with some representation of each element of the stream, specifically a partial digest.
 
-		The prover (server) on the other hand must act somewhat as a database.
+		The prover (server) on the other hand will act as a database.
 		Internally, the prover maintains a representation of the stream elements witnessed such that it can answer queries from the verifier.
 		That internal representation is the hash tree mentioned earlier which allows for a node to be update without rehashing its children.
-		The verifier exploits this property to maintain and update the root digest of the tree without communication with the prover.
+		The verifier will exploit this property to maintain and update the root digest of the tree without communication with the prover.
 
@@
 
-		During or after the streaming phase, the verifier may wish to query the prover (server) for the membership, successor, range, frequency etc of a universe element that maybe or may not have been witnessed.
-		The prover is expected to provide and answer to the query as well as a proof that can be efficiently verified by the verifier.
-		Figure \ref{fig:proof} in a visual representation of a membership proof for the green node.
-		The set of green nodes comprises the \textit{update path} for the green leaf node
-		The set of blue nodes are the \textit{siblings} of the nodes on the update path.
-		The prover then responds to a membership query with the computed answer as well as the set of sibling pairs which the verifier can then verify.
+		During or after the streaming phase, the verifier may wish to issue queries to the prover (server) of the form (non-)membership, successor, range, frequency etc. 
+		The prover is expected to provide an answer to the query as well as a proof that can be efficiently verified by the verifier.
+		Figure \ref{fig:proof} gives a visual representation of a membership proof for the green leaf node.
+		The whole set of green nodes comprises the \textit{update path} for the green leaf, while
+		the set of blue nodes is the \textit{siblings} of the nodes on the update path.
+		The prover would respond to a membership query with the computed answer as well as the set of labels of the sibling pairs as proof for the query answer.
 
-		Verification is done in two parts.
-		First, for each sibling pair is hashed to obtain a digest of their parent.
-		That calculated digest must be an equivalent representation of the label in the next pair representing this pairs parent.
-		Lastly, the digest produced by hashing the last pair of children must be equivalent to the running digest the verifier has been maintaining.
-		Shi and Papamanthou \cite{dads} describe the correctness properties in much more detail.
+		Verification is then done by the verifier in two parts.
+		First, each sibling pair is hashed to obtain a digest of their parent.
+		That calculated digest must be an equivalent representation of this pair's parent label, obtained from the proof.
+		Lastly, the final digest produced by hashing the last pair must be equivalent to the running digest the verifier has been maintaining.
+		Shi and Papamanthou \cite{dads} describe the correctness properties in sufficient detail.
 
@@
 	Implementing the SADS scheme was organized as follows.
-	First to be implemented was the data underlying structure used by the prover and mechanisms used by both prover and verifier to support the SADS scheme.
-	Then the data structures and mechanism to support the query phase was added.
-	Finally, these two parts were packaged such that another project may leverage the SADS scheme as a library primitive and build applications.
+	The first component to be implemented was the underlying datastructure used by the prover and the mechanisms used by both prover and verifier to support the SADS scheme; mechanisms for determining a node's index in the tree and calculating partial digests for example.
+	Then the data structures and mechanisms to support the query phase was added.
+	Finally, these two parts were packaged as a library such that another project may leverage the scheme primitive and build applications on top of SADS.
 
 	Our language of choice for this first stage of the project was Ruby.
-	Ruby's main advantage was the ease of rapid prototyping as well as the vast collection of libraries to enhance development and functionality available.
+	Ruby's main advantage was the ease of rapid prototyping as well as the vast collection of libraries available in the community.
 	Specifically, the main component of the SADS scheme is the hash function which require matrix multiplication support.
 	Ruby's standard library includes Vector and Matrix arithmetic operations which were sufficient for prototype scale implementations.
-	Ruby is also a forward looking choice as we still have the option to create bindings for C and C++ code, should we require more efficient linear algebra libraries.
+	Ruby was also a forward looking choice as we still have the option to create bindings for C and C++ code, when we require more efficient linear algebra libraries.
 
@@
 	The classes are, not surprising, Prover and Verifier and the module is named Sads.
-	Ruby has a feature called \textit{mixins} in which a module can be "mixed into" a class.
-	Mixing in a module has the effect of include all the modules fields and methods into the class definition.
-	The Sads module contains any pieces of functionality required by both the Prover and the Verifier - the definition of the hash function for example.
-	However since an instantiation of a Sads would not make sense, module mix in is used instead of class inheritance.
-	The Prover and Verifier classes then mix in the functionality defined in Sads augmenting the Prover-only or Verifier-only functionality defined in their respective classes.
-		Future iterations of this library will include a stronger focus on object-oriented design especially surrounding the nodes of the tree stored by the prover, the proofs generated per query requests, and abstractions surrounding the matrix multiplication so various linear algebra implementations can be tested.
+	Ruby has a feature called \textit{mixins} in which a module can be ``mixed into" a class.
+	Mixing in a module has the effect of include all the module's fields and methods into the class definition.
+	The Sads module contains any pieces of functionality required by both the Prover and the Verifier; the definition of the hash function for instance.
+	However, since an instantiation of a Sads would not make sense, module mix-in was used instead of class inheritance.
+	The Prover and Verifier classes then mix-in the functionality defined in Sads augmenting the Prover-only or Verifier-only functionality defined in their respective classes.
+		Future iterations of this library will include a stronger focus on object-oriented design especially surrounding the nodes of the tree stored by the prover, the proofs generated per query requests, and abstractions surrounding the matrix multiplication so various linear algebra implementations can be swapped in and tested.
 

### 2013-03-29 doc/final_report/draft_report.tex

@@
 	\end{equation}
-	The key to the SADS scheme is Merkle tree using a homomorphic hash function which allows 
-	In the case of the SADS scheme, the hash used is one with homomorphic properties allowing each node in the tree to be represented as a sum of parts, including the root of the tree.
+	The key to the SADS scheme is Merkle tree using a homomorphic hash function which allows each node in the tree to be represented as a sum of parts, including the root of the tree.
 
@@
 		Lastly, the final digest produced by hashing the last pair must be equivalent to the running digest the verifier has been maintaining.
-		Shi and Papamanthou \cite{dads} describe the correctness properties in sufficient detail.
+		Shi and Papamanthou \cite{sads} describe the correctness properties in sufficient detail.
 
@@
 	Implementing the SADS scheme was organized as follows.
-	The first component to be implemented was the underlying datastructure used by the prover and the mechanisms used by both prover and verifier to support the SADS scheme; mechanisms for determining a node's index in the tree and calculating partial digests for example.
+	The first component to be implemented was the underlying data structure used by the prover and the mechanisms used by both prover and verifier to support the SADS scheme; mechanisms for determining a node's index in the tree and calculating partial digests for example.
 	Then the data structures and mechanisms to support the query phase was added.
@@
 
-	The next stage was to implement streaming setting in which the Prover and Verifier operate independently while witnessing a common stream of data.
+	The next stage was to implement the streaming setting in which the Prover and Verifier operate independently while witnessing a common stream of data.
 	We wrote a very simple script to simulate such a scenario.
-	This script was created as a new project, also using Ruby, and installed the SADS library via the RubyGems command 'gem install --local'.
-	We wanted to ensure that the steps we took to use the SADS library in another project were the same or close to what other developers might do when using our library.
+	This script was created as a new project, also using Ruby, and installed the SADS library via the RubyGems command `gem install --local $<$path to library$>$'.
+	We wanted to ensure that the steps we took to use the SADS library in another project were the same as what we would expect other developers to do when using our library.
 	The script takes three parameters to initialize the entire protocol, the security parameter, the upper bound on the size of the stream, and the size of the universe of stream elements.
@@
 	The Verifier is then initialized with parameters provided to it from the Prover, similar to a client-server initialization wherein a client requests a session from a web server and can only proceed once the server grants the client the credentials.
-	Finally in this small example, the user is then able to run commands such as "add" which simulates the Prover and Verifier both witnessing some stream element or "query" which simulates a membership query from the Verifier to the Prover.
+	Finally in this small example, the user is then able to run commands such as ``add" which simulates the Prover and Verifier both witnessing some stream element or "query" which simulates a full membership query and response verification between Verifier to the Prover.
 
-	The final stage is to implement the simulation across distributed computing infrastructure, that is a Prover and Verifier non-co-located.
+	The final stage is to implement the simulation across distributed computing infrastructure, that is, with a Prover and a Verifier which are not co-located.
 	We imagine at least two possibilities for this implementation.
-	One possibility could be a web service could be created that plays matchmaker between "needy" clients requesting prover services and "generous" clients wishing to act as a prover.
+	One possibility could be a web service that plays matchmaker between ``needy" clients requesting prover services and ``generous" clients wishing to act as a prover.
+	The web service would then facilitate communication between the two parties but would not provide the computational resources to act as either the prover or verifier.
 	Another possibility, similar to the first, is one in which the web service itself acts as a Prover for any client requesting service.
-	In either of these scenarios we imagine deploying a server application possibly build in EventMachine or Node.js or some similar framework.
+	In either of these scenarios we imagine deploying a server application possibly built with EventMachine or Node.js or some other similar web application framework.
 	Some proposed applications include monitoring and verifying of bandwidth charges from an ISP or continuous sensor data.
@@
 \label{sec:forward}
-	\subsection{ Production Quality Libraries }
-
-	Through our benchmarking and profiling efforts, we hope to identify and fix the bottlenecks in implementation resulting in faster and higher quality code.
-	We recognize that the built-in implementations of certain high cost operations, such as matrix multiplication may end up being the limiting factor, and as such we've identified a few options for future optimization.
-	We hesitate to invest time and effort into any of these options too soon as we know that our first-draft implementation necessarily has room for improvement, but we outline those options here.
-
-	\begin{enumerate}
-	\item \textbf{SciRuby} : A Ruby library providing efficient implementations of linear algebra operations as well as other science, engineering, and visualization packages.
-\texttt{http://sciruby.com/}
-	\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length integers, for vectors, matrices and polynomials over the integers and over the finite fields and for arbitrary precision floating point arithmetic.
-\texttt{http://shoup.net/ntl/}
-	\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of the important concrete classes of structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, module theory and the theory of algebras.
-	In addition, certain families of structures from algebraic geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
-	\end{enumerate}
 
@@
 
+	\subsection{ Production Quality Libraries }
+
+	Through our benchmarking and profiling efforts, we hope to identify and fix the bottlenecks in implementation resulting in faster and higher quality code.
+	We recognize that the built-in implementations of certain high cost operations, such as matrix multiplication may end up being the limiting factor, and as such we've identified a few options for future optimization.
+	We hesitate to invest time and effort into any of these options too soon as we know that our first-draft implementation necessarily has room for improvement, but we outline those options here.
+
+	\begin{enumerate}
+	\item \textbf{SciRuby} : A Ruby library providing efficient implementations of linear algebra operations as well as other science, engineering, and visualization packages.
+\texttt{http://sciruby.com/}
+	\item \textbf{NTL 5.5.2} : A free software written in C++ providing data structures and algorithms for arbitrary length integers, for vectors, matrices and polynomials over the integers and over the finite fields and for arbitrary precision floating point arithmetic.
+\texttt{http://shoup.net/ntl/}
+	\item \textbf{MAGMA V2.18} : The kernel of Magma contains implementations of many of the important concrete classes of structure in five fundamental branches of algebra, namely group theory, ring theory, field theory, module theory and the theory of algebras.
+	In addition, certain families of structures from algebraic geometry and finite incidence geometry are included. \texttt{http://magma.maths.usyd.edu.au/}
+	\end{enumerate}
+
 	%\subsection{Semester Outlook}
